
```{r bloomr-init, include=FALSE}
## uncomment for BloomR distro
#library("rJava")
#library("Rbbg")

##Common packages
#library(zoo)
#library(xts)
#library(stringr)

```


# BloomR facility functions


bbg.bulk.tiks
=============
*Bulk historical data*  
Returns the historical data for a vector of tickers in xts or list format

Usage
------
    bbg.bulk.tiks(con, tiks, start=Sys.Date()-5, field="PX_LAST",
		addtype=FALSE, showtype=FALSE, use.xts=TRUE,
		price=TRUE, nrow=5, empty.sec=0) 
  
Arguments
---------
tiks
:   character vector of the tickers queried for data  

For other arguments see the function `bbg.bulk.csv` 

Details
--------
If an element of `tiks` is `NA` or empty (`""`) it is ignored. This is intended to avoid errors when the cahracter vector are read from a CSV file with empty cells. 

If `con=NULL` values are simulated by means of `bbg.sample()`. Sampled values are based on default values of `bbg.sample()`, but it is possible to set explicitly  `empty.sec, start, nrow`;  `sec.names` depends on `tiks` argument. See `bbg.sample()` help for more.

Value
------
If `use.xts=TRUE`, an xts object, where each column is the historical data of a security.  
If `use.xts=FALSE`, a list, where each element is the historical data of a security.


```{r bbg.bulk.tiks, include=FALSE}
bbg.bulk.tiks=function(
    con,  
    tiks, 
    start=Sys.Date()-5, field="PX_LAST",
    addtype=FALSE, showtype=FALSE, use.xts=TRUE,
    price=TRUE, nrow=5, empty.sec=0
    )
{

    ## Check connection
    if(!is.null(con) && !.bbg.is.con(con)) stop('Invalid connection parameter') 
    
    ## Check tickers (skip possible empty CSV cells)
    if(!is.character(tiks)) stop('Tickers should be in the form of a character vector')
    tiks=tiks[!is.na(tiks)] 
    tiks=tiks[tiks!='']
    
    ## Check start date
    if(is.na(as.Date(start, format='%Y%m%d'))) stop(paste('Invalida date', start))

    ## Check security type to add/show
    .bbg.check.type(addtype)
    .bbg.check.type(showtype)
    if(addtype==TRUE)  addtype="Equity"
    if(addtype!=FALSE) tiks=paste(tiks, addtype)
    if(!showtype) tiks.show=.bbg.cuttype(tiks) else tiks.show=tiks
    
    ## Check xts library availability
    if(use.xts && !require("xts", quietly=TRUE, character.only=TRUE))  stop("Can't find library xts")

    ## Get data as an xts class
    if(use.xts){ 
        LL = lapply(tiks, function(tik){
            cat('Loading', tik,  '\n')
            if(!is.null(con)) x=bdh(con, tik, field, start) else {
                x=bbg.sample(nrow, 1, price=price, start=start,
                    df=TRUE, sec.names=c('date', field))
            }
            x=xts(x[-1], as.Date (x[[1]]))
            if(nrow(x)==0) x=NA else x
        })

        ## Randomly identify empty.sec
    ##:ess-bp-start::browser@nil:##
browser(expr=is.null(.ESSBP.[["@4@"]]))##:ess-bp-end:##
        if(is.null(con)){
            x=round(length(tiks.show) * empty.sec) 
            empty=sample(length(tiks.show), x)
            LL[empty] = NA 
        }
        
        ## If there are only NAs cells output an empty xts else merge and fill empty cells with NAs 
        if(all(is.na(LL))) LL=na.omit(xts(t(unlist(LL)), Sys.Date())) else {
            if(length(LL)>1) LL=do.call("merge.xts", LL) else
            LL=LL[[1]]
        }
    
        ## Set labels
        if(nrow(LL)>0) names(LL) = tiks.show  else  dimnames(LL)= list(NULL,tiks.show)
        LL
        
    } else {
        ## Get data in list format
        LL=lapply(tiks, function(tik){
            cat('Loading', tik,  '\n')
            if(!is.null(con)) bdh(con, tik, field, start) else {
                bbg.sample(nrow, 1, price=price, empty.sec=empty.sec, start=start,
                           df=TRUE, sec.names=c('date', field))
            }            
        })
        
        ## Randomly identify empty.sec
        if(is.null(con)){
            x=round(length(tiks.show) * empty.sec) 
            empty=sample(length(tiks.show), x)
            LL[empty] = NA 
        }
        
        setNames(LL, tiks.show)
    }
}
```


bbg.bulk.csv
=============
*Historical from grouped tickers in a CSV files*  
Reads a CSV file containing a group of tickers in each column and returns the historical data in xts or list format. The CSV file is assumed to have headers denoting group labels. 


Usage
------
    bbg.bulk.csv(con, file, start = Sys.Date() - 5, field = "PX_LAST", 
        cols = NULL, addtype = FALSE, showtype = FALSE, use.xts = TRUE, 
        comma = TRUE,
	    price=TRUE, nrow=5, empty.sec=0
		) 
  
Arguments
----------
con
:   the connection token returned from bbg.open(). If `NULL` simulated values are generated.   
file
:   path to CSV file.  
start
:   start date. Can be a Date object or an ISO string without separators. Defaults to 5 days before current date.  
field
:   String denoting the Bloomberg field queried. Defaults to "PX_LAST". If the field is wrong or not accessible, data will be empty but no error will be raised.  
cols
:   Logical or integer vector for selecting CSV columns (ticker groups). Defaults to all columns.  
addtype
:   If a string denoting the security type, it will be added to all tickers; if TRUE "Equity", will be added; if FALSE (the default), nothing will be added.  
showtype
:   if TRUE, security types will be removed from names of list or xts output. It defaults to FALSE.  
use.xts
:   if TRUE (the default) each group will be formatted as an xts object else as a list.  
comma
:   to be set to FALSE for (non-English) CSV, using semicolon as separator.  
nrow
:   Maximum number of simulated rows (actual is random). Ignored if `con!=NULL`, it defaults to 5.
empty.sec
:   ratio of securities returning no data. Ignored if `con!=NULL`, it defaults to 0.

Details
-------
Empty CSV cells or cells interpreted as NAs will be ignored.  
If `con=NULL` values are simulated by means of `bbg.sample()`. This function is used with default values, except for `empty.sec, start, nrow`, which can be explicitly passed as arguments, and `sec.names` depending on on tickers found in the CSV file. See `bbg.sample()` help for more.

Value
------
a list where each element is the historical data of a CSV group.  
If `use.xts=TRUE`, elements are xts object, where each column is the historical data of a security.  
If `use.xts=FALSE`, elements are sub-list, where each element is the historical data of a security.  
If there is only one group, the first (and unique) element of the list will be returned.  


```{r bbg.bulk.csv, include=FALSE}
bbg.bulk.csv=function(con, file, start=Sys.Date()-5, field="PX_LAST", cols=NULL,
    addtype=FALSE, showtype=FALSE, use.xts=TRUE, comma=TRUE,
    price=TRUE, nrow=5, empty.sec=0
    )
{

    ## Check csv file
    if(!file.exists(file)) stop(paste('Unable to find the csv file', file))
    if(comma) csv=read.csv(file=file, as.is=TRUE) else{
        csv=read.csv2(file=file, as.is=TRUE)}

    ## Check cols arg
    if(!is.null(cols)){
        if(is.logical(cols) && !length(cols)==ncol(csv))
            stop(paste('Length of logical vector', paste(cols, collapse=' '),
                       'not equal to number of groups in', file))

        if(is.integer(cols) && max(cols)>ncol(csv))
            stop(paste('Unable to subset groups in', file, 'with columns', paste(cols, collapse=' ')))

        if(!is.logical(cols) && !all(cols%%1==0)) stop(paste(
         "'Col argument should be an integer or a logical vector of the same length of the groups in", file))
        csv=csv[cols]        
    }

    ## Get group names and count
    gnams=names(csv)
    gcnt=ncol(csv)

    ## Loop groups in csv
    grps=list()
    for(g in 1:gcnt){
        cat('Processing', gnams[g],  '\n')
        x=list(bbg.bulk.tiks(con, csv[[g]],
            start, field, addtype, showtype, use.xts, price=price, nrow=nrow, empty.sec=empty.sec))
        names(x)=gnams[g]
        grps=c(grps, x)        
    }
    if(length(grps)==1) grps=grps[[1]]
    grps
}
```

bbg.bulk.idx
============

Description
-----------
Returns the historical data for the constituents of an index in xts or list format.

Usage
-----
    bbg.bulk.idx(con, index, start=Sys.Date()-5, field="PX_LAST",  
        include.idx=TRUE, use.xts=TRUE)

Arguments
---------
index
:   string denoting the index ticker with or without the final security type label ('Index')  
include.idx
:   if TRUE (default) returns also historical data for the index.  
For other arguments see the function `bbg.bulk.csv`  

Value
-----
If `use.xts=TRUE`, an xts object, where each column is the historical data of a constituent.  
If `use.xts=FALSE`, a list, where each element is the historical data of a constituent.  
If `include.idx=TRUE`, the last column or element will be the historical data of the index.  


```{r bbg.bulk.idx, include=FALSE}
bbg.bulk.idx=function(con, index, start=Sys.Date()-5, field="PX_LAST",
    include.idx=TRUE, use.xts=TRUE)
{

    ## Check connection
    if(!.bbg.is.con(con)) stop('Invalid connection parameter') 

    ## Check index format. Add 'INDEX' if missing
    if(!is.character(index)) stop('Index should be a string')
    if(length(index)>1) stop('Only one index')
    if(!grepl("INDEX$", toupper(index))) index=paste(index, 'INDEX') 

    ## Get index members
    tiks=bds(con, index, 'INDX_MEMBERS')
    tiks=paste(tiks[[1]], 'Equity')

    ## Include index?
    if(include.idx) tiks=c(tiks, index)

    ## Get data
    bbg.bulk.tiks(con, tiks, start, field, addtype=NULL, use.xts=use.xts) 

}
```

bbg.desc
========

Description
-----------
Get security descriptions.

Usage
-----
    bbg.desc(con, tik)

Arguments
---------
con
:   the connection token returned from bbg.open()
tik
:   string denoting the ticker queried for data 

Value
-----
A data frame containing the value of the Bloomberg fields form `ds001` to `ds009` and the long field `CIE_DES_BULK`.

```{r bbg.desc, include=FALSE}
bbg.desc=function(con, tik)
{

    ## Check connection
    if(!.bbg.is.con(con)) stop('Invalid connection parameter') 

    ## Check ticker format
    if(!is.character(tik)) stop('The ticker should be a string')
    if(length(tik)>1) stop('Only one  ticker')
       
    ## Short description fields as data frame
    des=paste0('ds00', 1:9)
    des=des[-7] # not usually working 
    x=bdp(con, tik, des)
    x=data.frame(t(x), stringsAsFactors=FALSE)

    ## Long description field
    xx=bds(con, tik, 'CIE_DES_BULK')

    ## Merge fields add long desc to DF
    if(!is.null(xx)) colnames(xx) = colnames(x)
    rnams=c(rownames(x), rownames(xx))
    x=rbind(x,xx)
    rownames(x)=rnams
    x
}
```


bbg.bulk.desc
=============

Description
-----------
Get security descriptions for a vector of tickers.

Usage
-----
    bbg.bulk.desc(con, tiks) 

Arguments
---------
con
:    the connection token returned from bbg.open()
tiks
:    character vector of the tickers queried for data

Value
-----
A list of data frames, each representing the description of a security. For the format of data frames see the function `bbg.desc`.


```{r bbg.bulk.desc, include=FALSE}
bbg.bulk.desc=function(con, tiks) {

    LL = lapply(tiks, function(tik){
        cat('Reading', tik,  '\n')
        bbg.desc(con, tik)             
    })
    names(LL)=tiks
    LL
}

```


bbg.sample
==========

Description
------------
Return simulated historical data for n securities in xts or df format.

Usage
-----
    bbg.sample(nrow, nsec=1, price=TRUE, start=Sys.Date(), mean=ifelse(price, 10, 0.1), sd=1,
       jitter=0, same.dates=FALSE, no.na=FALSE, empty.sec=0, df=FALSE, sec.names=NULL)

Arguments
---------
nrow
:   number of simulated data points for each security; if `same.dates=FALSE`, the number of rows for each sampled security will be a random number not exceeding nrow, else it will be nrow for all securities.
nsec
:   number of simulated securities (defaults to 1).
price
:   if TRUE (default), simulated values are non-negative.
start
:   start date. Can be a Date object or an ISO string without separators. Defaults to current date.
mean
:   mean of security generated values. If `price=TRUE`, default to 10 else defaults to 0.1.
sd
:   sd of security generated values. It defaults to 1.
jitter
:   modifies each security mean by adding adding a random value in [-jitter, jitter]. Defaults to 0.
same.dates
:   if TRUE, all sampled securities will refer to the same dates and for each security the number will equal nrow. If FALSE (default), date values and number will randomly differ. For each security the random number will not exceed `nrow`.
no.na
:   if `same.dates=FALSE`, when merging sampled security data NAs are likely to be produced. If `no.na=FALSE` (default) they will be left, otherwise they will be removed using R `na.omit`
df
:   if FALSE (default), the output will be an xts object, else the output will be a data frame with the first column containing the dates of the sampled data.
sec.names
:   character vector for column names. If `df=FALSE` the length of the vector should be equal to `nsec`, else to `nsec + 1` (because of the first column containing dates). By default security names are like 'sample1', 'sample2', etc. and the date column is named 'date'.
empty.sec
:   ratio of securities returning no data (defaults to 0). The result is rounded without decimal places.

Value
-----
If `df=TRUE`, a data frame object, where the first column is the vector with all generated dates merged and each  subsequent column contains the sampled data of a security. If `df=FALSE`, an xts object, where each element is the sampled data of a security, while the dates will be part of the xts time object. In both cases if `same.dates=FALSE` and/or `empty.sec!=0` generated data points will have different length and the the date gaps will be filled with NAs, except if `no.na=TRUE`. If the generated values are only NAs the output will be converted to a 0-rows xts or data frame, containing only security labels accessible with `dimnames(*)[[2]]`. 

```{r bbg.sample, include=FALSE}
bbg.sample=function(nrow, nsec=1, price=TRUE, start=Sys.Date(), mean=ifelse(price, 10, 0.1), sd=1,
    jitter=0, same.dates=FALSE, no.na=FALSE, df=FALSE, empty.sec=0,sec.names=NULL)
{
 
    if(!require("xts", quietly=TRUE, character.only=TRUE))  stop("Can't find library xts")

    ## Start can be Date class or ISO string without sep
    if(is.na(as.Date(start, format='%Y%m%d'))) stop(paste('Invalida date', start)) else 
        start=as.Date(start, '%Y%m%d')

    ## Check sec.names
    if(!is.null(sec.names)) {
        if(!is.character(sec.names)) stop("'sec.names' should be a character vector")
        if(!df && length(sec.names)!=nsec) stop("'sec.names' length should be equal to 'nsec'")
        if(df && length(sec.names)!=nsec+1) stop("'sec.names' length should be equal to 'nsec'+1")
    }

    ## Check empty.sec is a ratio
    if(empty.sec<0 || empty.sec>1) stop("'empty.sec' must be between 0 and 1")

    ## Randomly identify empty.sec
    x=round(nsec * empty.sec) 
    empty=sample(nsec, x)
    
    ## Make xts matrix    
    tss=lapply(1:nsec, function(col){

        ## Jitter 
        mean.jit= mean + runif(1, -jitter, jitter)
    
        ## Generate TS with fixed or random dates
        if(same.dates) r=nrow else r=sample(1:nrow,1)

        ## Generate column if not among the empty ones
        if(col %in% empty){
            x=NA
        } else {            
            x=xts(round(rnorm(r,mean.jit, sd),3), sort(sample(Sys.Date()+1:nrow,r)))            
            if(price) coredata(x)=abs(coredata(x))   # Price always non-negative
        }
                
        x
    })

    ## If there are only NAs cells output an empty date/val xts else merge and fill empty cells with NAs 
    if(all(is.na(tss))) tss=na.omit(xts(t(unlist(tss)), Sys.Date())) else {
        if(length(tss)>1) tss=do.call("merge.xts", tss) else
        tss=tss[[1]]
    }

    ## Remove NAs
    if(no.na) tss=na.omit(tss)
    
    ## Set labels
    x=paste0('sample', 1:ncol(tss))
    if(nrow(tss)>0)  names(tss) = x  else  dimnames(tss)= list(NULL,x)
        
    ## Convert to data frame
    if(df) tss=data.frame(date=time(tss), tss)
    
    if(is.null(sec.names)) tss else
       setNames(tss, sec.names)
   
}
```



Internal bbg functions
=======================

Description:
------------
Returns the historical data for the constituents of an index in xts or list format.

Usage:
------

    .bbg.is.con(con)
    .bbg.types
    .bbg.check.type(type) 
    .bbg.cuttype(type)
    .bbg.jar()

Arguments:
----------
con
:   the connection token returned from bbg.open()
type
:   a string representing the security type

Details
-------
`.bbg.is.con` checks for the validity of a connection token.
`.bbg.types` is a character vector with security types suitable as an argument for `bbg.bulk*` functions.
`.bbg.check.type` checks if a type matches `.bbg.types`.
`.bbg.cuttype` cuts trailing security type from character vector.
`.bbg.jar()` returns the path to the blpapi*.jar


```{r bbg-internal, include=FALSE}

## Check connection token
.bbg.is.con=function(con) identical(attr(con, 'jclass'), "org/findata/blpwrapper/Connection")

## Legal security types
.bbg.types=c('Govt', 'Corp', 'Mtge', 'M-Mkt', 'Muni', 'Pfd', 'Equity', 'Comdty', 'Index', 'Curncy')

## Check security type
.bbg.check.type=function(type) {
    if(is.character(type)){
	x=toupper(type)
	xx=toupper(.bbg.types)
	if(!any(xx %in% x)) stop(paste(x, 'not in', paste(xx, collapse=' ')))
    }
}

## Cut trailing security type from character vector 
.bbg.cuttype=function(type){
    p=paste0(' +', .bbg.types, '$|', collapse='')
    p=sub('\\|$', '', p)
    sub(p, '', type, ignore.case=TRUE)
}


.bbg.jar=function(){
	jarpath=paste0(R.home(), "/blpapi_java/bin")
        Sys.glob(file.path(jarpath,  "blpapi-[0-9]*.jar"))
    }

```

Manage connections
==================

Description
------------
Open and close the connection to the Bloomberg service. 

Usage
-----
    bbg.open()
    bbg.close(con)
 
Arguments
---------
con
:   the connection token returned from bbg.open()

Example
-------

    con=bbg.open() ## Open the connection and get the token
    ## Get market data
    bbg.close(con) ## Use the token to release the connection



```{r connections, include=FALSE}

bbg.open=function() blpConnect(blpapi.jar.file=.bbg.jar())
bbg.close=function(conn)  blpDisconnect(conn)

## misc func
delete.all= function() rm(list=ls(all=TRUE))

```

Time extension functions
=========================

Description
------------
Functions to get, set dates.

Usage
-----
    day(d)
    month(d)
    year(d)
    day(d, n)
    month(d, n)
    year(d, n)
    day(d)=x
    month(d)=x
    year(d)=x
    d %+% n
    d %-% n
    last.day(d)
    day.us(d1, d2)
    delete.all()

Arguments
---------
d, d1, d2
:   objects of class date
x
:   an integer representing the day/month/year
n
:   an integer representing the months to add/subtract


Details
-------
If `component` is `day`, `month` or `year`: `component(d)` returns the *component* of the date `d` as an integer; `component(d, n)` returns the date `d` with the *component* set to the integer `n`; `component(d)= n` sets to the *component* of the date `d` to the integer `n`.  
`%+%` and `%-%` add and subtract months to a date.  
`last.day` returns last day of the month as an integer. `day.us` calculates date differences with the US convention.  
`delete.all` deletes all objects (variables) from memory. 


```{r time, include=FALSE}
`%+%` <- function(x,y) UseMethod("%+%")
`%+%.Date` <- function(date,n) seq(date, by = paste (n, "months"), length = 2)[2]
`%-%` <- function(x,y) UseMethod("%-%")
`%-%.Date` <- function(date,n) seq(date, by = paste (-n, "months"), length = 2)[2]

year=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%Y")) else year(d)=n
    d    
}
`year<-`=function (d, value) d <-as.Date(paste0(value, format(d, "-%m-%d")))

month=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%m")) else month(d)=n
    d 
}
`month<-`=function (d, value) d <-as.Date(paste0(format(d, "%Y-"),  value, format(d, "-%d")))

day=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%d")) else day(d)=n
    d 
}
`day<-`=function (d, value) d <-as.Date(paste0(format(d, "%Y-%m-"),  value))

last.day=function(d){
    x=d %+% 1 #add a month
    day(x)=1  #set to 1st
    day(x-1)  #get day before
}

day.us=function(d1, d2){
    #set to first of month
    x1=day.mod(d1,1);x2=day.mod(d2,1);
    x=seq(x1, x2, by="1 month")
    #last day of each month in seq
    x=sapply(x, last.day)
    #count 31d-months
    x=length(which(x>30))
    #substract 1 for each 31d-month
    as.numeric(d2-d1-x)
}

## misc func
delete.all= function() rm(list=ls(all=TRUE))

```

```{r unitests, eval=FALSE, include=FALSE}

#Unit tests
#=========

## pars

file='bric.csv'
start="20060101"
field="PX_LAST"
cols=NULL
cols=1
addtype=NULL
addtype='Equity'


#con=bbg.open()
# 
#start="20140301"
#bric.focus=
#    bbg.bulk.csv(con, file, field="PX_LAST", cols=c(1), use.xts=TRUE)
# 
#save(bric.focus, file='bricFocus.RData', compress='xz')
#save(bric.focus, file='bricFocus.RData')
# 
#bbg.close(con)
# 
#x=setNames(bbg.sample(4,1), field); data.frame(date=time(x), x)
#bbg.bulk.tiks(con,  lab, start=Sys.Date()-5, field="PX_LAST", addtype=TRUE)
#bbg.bulk.desc(con, c("MSFT US EQUITY", "ASSGEN 10.125 07/10/2042 Corp")) 
# bbg.bulk.idx(con, "SX5E Index", start=Sys.Date()-5, field="PX_LAST")


```
