---
title: "My First Report"
---

Introduction
============

Here are some formatting examples. See the text of the Rmd file to understand them:

**Bold**, *Italic*, _Italic again_

These lines will be
printed as a single line.

To insert a line break leave two or more blank spaces to the right of this line.
Now I will be printed on a new line!

The following is the typewriter style. Use it if you want to comment/show code templates.

    # This is the main project function
    myfunction(arg1, arg2, arg3)

To obtain it leave four spaces before lines and (at least) one blank line above and below the block.
To use the typewriter style *inline*, enclose the text inside backticks,
therefore this `text` will be printed in typewriter style.
If you use dashes, `-`, instead of equals, `=`, you will get a subsection header.

Using Itemisations and Enumerations
-----------------------------------

See code Rmd file content

* Item 1
* Item 2
* ...

1. Item 1
2. Item 2
3. ...


And now some R code 
-------------------

```{r setup, echo=FALSE}
# set global chunk options: images will be 7x5 inches
opts_chunk$set(fig.width=7, fig.height=5)
```

I will implement some calculations. See the code and results below:

```{r myFirstChunk}
x <- 1+1 # a simple calculator
set.seed(123)
rnorm(5)  # generate five random numbers
```

I will now generate six random numbers.
Well, you already know the code behind, so I don't show it:

```{r myChunkNoecho, echo=FALSE}
set.seed(321)
rnorm(6)  # generate six random numbers
```

And now let me plot my ideas:

```{r myNicePlot, fig.cap="In this plot we compare Miles per Gallon with HP."}
par(mar = c(4, 4, 2, .1))
with(mtcars, {
  plot(mpg~hp, pch=20, col='darkgray', main="My First Plot")
  lines(lowess(hp, mpg))
})
```

Maths
-----


You can use LaTeX style maths, therefore:

    $$f(\alpha) = \frac{1}{c+1}\sum_{k=0}^\infty  x^{\alpha-1}_k,
                  \mathrm{ where } c \in \mathbf{n} = \{1 \ldots \bar{n} \}$$

will be rendered as:

$$f(\alpha) = \frac{1}{c+1}\sum_{k=0}^\infty  x^{\alpha-1}_k,
              \mathrm{where} c \in \mathbf{n} = \{1 \ldots \bar{n} \}$$.


For inline expressions use a single dollar. So `$x_k^j$` gives: $x_k^j$.

Note that `_` and `^` set indices and exponents. LaTeX commands/symbols are preceeded by a backslash, e.g.: `\frac`, `\sum`, and arguments (if any) are encolesed in curly braces. To use braces literally escape them with : `{` and `\}`. A single index or exponent does not require enclosing braces.  
Exercise on some online engine like  [codecogs.com](https://www.codecogs.com/latex/eqneditor.php).


Presenting matrix-like objects as tables
---------------------------------------

To print matrix-like objects as tables, you use the command `kable` and the chunk option `results="asis"`.

First of all, normally you don't get table captions using Pandoc.
But there is a trick to get them: add a chunk on top of your `my-first-report.Rmd` with the usual _chunk-syntax_  containing the following:

```{r setTableCaptions, eval=FALSE}
opts_knit$set(rmarkdown.pandoc.to = "latex")  
```

Add a chunk for creating a generic matrix to print:

```{r mymat}
i=4; j=6
(M=matrix(rnorm(i*j), i, dimnames= list(letters[1:i], LETTERS[1:j])))
```

To print `M` with four decimals, we use this chunk: 

```{r myMat.print}
kable(M, digits=4, caption="This is a test matrix")
```

One can show a tabular regression output like follows:


```{r reg.print}
x=rnorm(100)
y=x^2
kable(summary(lm(y ~ x))$coefficients)
```

```{r compile, echo=FALSE, eval=FALSE}

## Warning sync work dir using: setwd(), alt-enter, the BRemacs sync button 
br.rmd2both("my-first-report.Rmd")

## HTML only for BloomR Core
## library(knitr)
## library(markdown)
## knit("my-first-report.Rmd")
## markdownToHTML("my-first-report.md", "my-first-report.html")

## Note the eval=FALSE above
## This is for internal commands that should not appear in PDF/HTML report

```
