---
title: "BloomR Rmd Functions"
author: "Antonio Fasano"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
---

```{r build, purl=FALSE, eval=FALSE, include=FALSE}

knit.twice <- function(this.file) {# knit this Rmd. 

    ## Polymode tends to start R synced with git root dir.
    ## We check the Rmd path is absolute or as expected    
    GITPOS <- "/src/br-libs"
    abspt <- function(pt) unlist(strsplit(pt, split="/"))[[1]] == "" || regexpr("^.:(/|\\\\)", pt) != -1L
    is.rel <- !abspt(this.file)
    bad.dir <- sub(paste0("^", dirname(dirname(getwd()))), "", getwd()) !=  GITPOS

    ## Try to fix a bad path, when not abolute
    if(bad.dir && is.rel) 
        if(dir.exists(file.path(normalizePath("."), "src/br-libs"))) {
            setwd("src/br-libs") # Polymode
        } else {
            stop("Rmd file expected to be in '", GITPOS, "'. The given absolute path is\n",
                 normalizePath(this.file), "\nUse an absolute path or check 'GITPOS' var in knit.twice()")
        }
            
    ## In BloomR use the shipped pandoc/LaTeX
    is.bloomr <- Sys.getenv("bloomr_branch") != ""
    wp <- function(expr) # Execute expression with modified PATH
        if(is.bloomr) .br.pathexe(substitute(expr), quiet = FALSE) else expr

    ## Prepare to knit
    knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60))
    fse <- tools:::file_path_sans_ext(this.file)

    ## Step 1 
    step2 <- FALSE
    knitr::knit(this.file)
    parseargs(paste0(fse, '.Rmd'), paste0(fse, '.tmp.Rmd'))
    
    ## Step 2 
    step2 <- TRUE
    knitr::knit(paste0(fse, '.tmp.Rmd'), output=paste0(fse, '.md'))
    topics(paste0(fse, '.Rmd'), paste0(fse, '.md'))
    wp(rmarkdown::render(paste0(fse, '.md'), "html_document"))
    wp(rmarkdown::render(paste0(fse, '.md'), rmarkdown::pdf_document()))
    unlink(paste0(fse, '.tmp.Rmd'))    
    file.rename(paste0(fse, '.tmp.R'), paste0(fse, '.R'))    

    ## Fix issue with purl commenting everything
    txt <- readLines(paste0(fse, ".R"))
    txt <- sapply(txt, function(line) sub("^#  ", "", line))
    writeLines(txt, paste0(fse, ".R"))

    ## Inform on new wdir
    if(bad.dir && is.rel) message("\nNOTE: Workdir was adjusted to ", GITPOS, ".")

}
knit.twice("bloomr-rmd.Rmd")

## if with markdown lib, but topics needs rewrite
## shell("pandoc bloomr.md -o bloomr.pdf", shell=Sys.getenv("COMSPEC"))
## markdownToHTML("bloomr.md", "bloomr.html") 
## markdownToHTML("README.md", "README.html")

```

# BloomR Rmd functions


```{r setup, purl=FALSE, include=FALSE}

### Chunk templates and defaults
knitr::opts_chunk$set(echo=TRUE)
OTS <- knitr::opts_template$set
## Step1 template: Load in memory and copy to <lib>.R
OTS(purlme=   list(purl=TRUE,  eval=!step2,   include=FALSE))
## Step1 template: Load in memory and copy to <lib>.R
OTS(demofull= list(purl=FALSE, eval=step2,  include=TRUE))  # run with code + output   
OTS(demorun=  list(purl=FALSE, eval=step2,  include=TRUE, echo=FALSE)) # run with output, but no code 
OTS(demohid=  list(purl=FALSE, eval=step2,  include=FALSE)) # run only (without code or output)
OTS(democode= list(purl=FALSE, eval=FALSE,  include=TRUE))  # show code only, no run  
OTS(buildenv= list(purl=FALSE, eval=step2,  include=FALSE)) # Called in step2 by the build chunk to knit this document 

## Purl hook (suggested in the past, but probably no necessary today)
knitr::knit_hooks$set(purl = knitr::hook_purl)

## Resource directory relative path
RESDIR <- "../../res"

```


```{r store, opts.label='purlme'}
## Purl this first
## Store br.* objects in dedicated namespace
bloomr.rmd <- new.env(parent=asNamespace("stats"))

## func: store(func);  var: store("var")
store=function(sym){
    if(is.function(sym)){
        name=deparse(substitute(sym)) 
        val=sym
    } else {
        name=sym
        val=get(sym)
    }
    
    assign(name, val, envir=bloomr.rmd)
    rm(list=name, envir=parent.frame())
}

```


```{r rmd-internal, opts.label='purlme'}
.br.addpaths <- function(pandonly = FALSE, quiet = TRUE){
### Add to Windows System Path the executable directories of LaTeX, Pandoc, and Perl with this search priority, and
### return invisibly the original path. If "pandonly" is true, add only Pandoc. If "quiet" is false, print the new path.

    ## Exit in case this snippet is not exectuted in Windows
    if(.Platform$OS.type != "windows")
        stop("Sorry, Bloomberg Terminal only exists for Windows and so BloomR functions.")
    
    ## Find executables
    xfind <- function(exepath){
        exe <- .br.home(exepath)
        if(!file.exists(exe))
            stop(paste("Unable to find:", exe, '\nYour BloomR edition might not support it.'))
        normalizePath(dirname(exe))
    }

    pandir <- xfind("pandoc/bin/pandoc.exe")
    if(!pandonly){
        latdir <- xfind("tinytex/bin/windows/latex.exe")    
        perldir <- xfind("tinytex/tlpkg/tlperl/bin/perl.exe")
        .br.inittex()
    }
    
    ## Add executable dirs to system path
    old.path <- Sys.getenv('Path')
    ap <- function(dir) Sys.setenv(PATH=paste0(dir, ";", Sys.getenv("PATH")))

    ## Wanted final order is: LaTeX;Pandoc;Perl 
    if(!pandonly) ap(perldir)
    ap(pandir)
    if(!pandonly) ap(latdir)
  
    ## Add Ghostscript environment variable   
    ## if(!pandonly)  Sys.setenv(GSC="mgs.exe") # only miktex?  

    if(!quiet) message(Sys.getenv("PATH"))

    invisible(old.path)
  
}
store(.br.addpaths)

.br.pathexe <- function(cmdexpr, pandonly = FALSE, quiet = TRUE){
### Change system path with .br.addpaths(), evaluate the expression cmdexpr in parent env, restotre path.
### See .br.addpaths() for pandonly and quiet arguments

    old.path <- Sys.getenv("PATH")
    tryCatch(
        {
            .br.addpaths(quiet=quiet)
            eval(cmdexpr, parent.frame())  
        },

        error = function(cond) {
            message(cond)
            return(NA)
        },

        warning = function(cond) {
            message(cond)
            return(NULL)
        }, 

        finally = {
            Sys.setenv(PATH = old.path)
        })
}
store(.br.pathexe)

.br.inittex <- function() {
    library('tinytex')
    options(tinytex.tlmgr.path = file.path(.br.home(), "tinytex/bin/windows/tlmgr.bat"))
}
store(.br.inittex)

```




br.md2pdf{#br.md2pdf} 
=====================

Description
-----------
Make a markdown file into a PDF
It assumes that you have installed the proper BloomR version.

Usage
-----
    br.md2pdf(md.file, pdf.file)

Arguments
---------
md.file
:   path to the markdown file to be converted.  

pdf.file
:   path to the PDF file to be generated. If missing, change extension of rmd.file to pdf.  

quiet
:   FALSE to show the system PATH variable.  

Details
-------
Stop with an error if you have not installed proper BloomR version.

Value
-----
If there are no errors, it returns zero invisibly, otherwise it prints an error message and returns the related error code.


```{r br.md2pdf, opts.label='purlme'}
br.md2pdf <- function(md.file, pdf.file, quiet=TRUE){
### Make a markdown file into a PDF
### You need the proper BloomR version

    ## Test arguments
    if(missing(md.file)) stop("Argument 'md.file' missing.")
    if(missing(pdf.file)) pdf.file=paste0(tools:::file_path_sans_ext(md.file), '.pdf')
        
    ## Set executable paths and render
    cmd <- paste("pandoc", .br.wpath(md.file), '-o', .br.wpath(pdf.file))
    cmdexpr <- quote(system(cmd, intern = TRUE, invisible = FALSE))
    out <- .br.pathexe(cmdexpr, quiet = quiet)
    invisible(out)
}
store(br.md2pdf)
```

br.rmd2html{#br.rmd2html} 
========================

Description
-----------
Make an R markdown file into a HTML.

Usage
-----
    br.rmd2html(rmd.file, html.file, quiet=TRUE)

Arguments
---------
rmd.file
:   path to the R markdown file to be converted.  

html.file
:   path to the HTML file to be generated. If missing, change extension of rmd.file to html.  

quiet
:   FALSE to show the system PATH variable and intermediate files.  

Details
-------
You need BloomR LaTeX addons or the proper BloomR version. 

Value
-----
If there are no errors, it returns invisibly the absolute path of the output file.


```{r br.rmd2html, opts.label='purlme'}
br.rmd2html <- function(rmd.file, html.file, quiet=TRUE){
### Make an R Markdown file into a HTML self-contained file
### You need the proper BloomR edition
    
    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    if(missing(html.file)) html.file=paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.html')
    if(html.file != basename(html.file)) stop("Sorry, currently you can only specify a name without directory as an output doc.")
    
    library(knitr)
    #library(rmarkdown)

    ## Render Rmd to HTML
    cmdexpr <- quote(
        rmarkdown::render(rmd.file,
               output_format = rmarkdown::html_document(theme="cerulean", highlight="tango",
                                           md_extensions="-tex_math_single_backslash"),
               output_file = html.file,
               clean = quiet, quiet = quiet)
    )    
    out <- .br.pathexe(cmdexpr, pandonly = TRUE, quiet = quiet)
    invisible(out)
}
store(br.rmd2html)
```
br.rmd2slides{#br.rmd2slides} 
============================

Description
-----------

`br.rmd2slides.html()` makes an R Markdown file into a [Google Slides](https://github.com/googleworkspace/md2googleslides) self-contained HTML file.  
`br.rmd2slides.pdf()` makes an R Markdown file into a Beamer PDF.  
`br.rmd2slides()` makes both a self-contained HTML file and a Beamer PDF.  
`highlight.styles()` print the list of available syntax highlight styles.


Usage
-----
	br.rmd2slides(rmd.file, theme = "AnnArbor", highlight = "tango", quiet = TRUE) 
    br.rmd2slides.html(rmd.file, html.file, quiet=TRUE)
    br.rmd2slides.pdf(rmd.file, pdf.file, theme = "AnnArbor", highlight = "tango", quiet = TRUE){
	highlight.styles()

Arguments
---------
rmd.file
:   path to the R markdown file to be converted.  

theme
:   A Beamer [theme](https://hartwork.org/beamer-theme-matrix/)  

highlight
:    A syntax highlight style. Use `highlight.styles()` to list them.  


html.file
:   path to the HTML file to be generated. If missing, we change the extension of `rmd.file` to ".html".  

pdf.file
:   path to the PDF file to be generated. If missing, we change the extension of `rmd.file` to ".pdf".  

quiet
:   FALSE to show the system PATH variable and intermediate files.  


Details
-------
You need the proper BloomR edition. The output filename for `br.rmd2slides()` is automatically generated by changing the `rmd.file`  file extension.

To represent the slide structure in an Rmd, you use the following pattern:

    ---
    title: "Your Presentation Title"
    author: "Your Name"
    date: "A date description"
    ---
     
    ## Slide 1
     
    This is the content for slide 1.
     
    ## Slide 2
     
    This is the content for slide 2.
     
    etc.

Then you can proceed as usual to add R code chunks.
 
Value
-----
If there are no errors, it returns invisibly the absolute path of the output file.


```{r br.rmd2slides, opts.label='purlme'}
br.rmd2slides.html <- function(rmd.file, html.file, quiet=TRUE){
### Make an R Markdown file into a Google Slides self-contained HTML file
### You need proper BloomR edition
    
    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    if(missing(html.file)) html.file=paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.html')
    if(html.file != basename(html.file)) stop("Sorry, currently you can only specify a name without directory as an output doc.")
    
    library(knitr)
    #library(rmarkdown)

    ## Render Rmd to HTML
    cmdexpr <- quote(
        rmarkdown::render(rmd.file,
               output_format = rmarkdown::ioslides_presentation(
                   md_extensions="-tex_math_single_backslash"
               ),
               output_file=html.file,
               clean=quiet, quiet=quiet)
    )
    out <- .br.pathexe(cmdexpr, pandonly = TRUE, quiet = quiet)    
    invisible(out)
}
store(br.rmd2slides.html)

br.rmd2slides.pdf <- function(rmd.file, pdf.file, theme = "AnnArbor", highlight = "tango", quiet = TRUE) {
### Make an R Markdown file into a Beamer PDF
### You need the proper BloomR edition

    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    if(missing(pdf.file)) pdf.file=paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.pdf')
    if(pdf.file != basename(pdf.file)) stop("Sorry, currently you can only specify a name without directory as an output doc.")

    library(knitr)
    #library(rmarkdown)

    ## Render Rmd to PDF
    cmdexpr <- quote(
        rmarkdown::render(rmd.file,
               output_format = rmarkdown::beamer_presentation(theme = theme, highlight = highlight,
                                          md_extensions="-tex_math_single_backslash",
                                          keep_tex = !quiet, keep_md = !quiet),
               output_file = pdf.file,
               clean = quiet, quiet = quiet)
    )

    out <- .br.pathexe(cmdexpr, quiet = quiet)
    invisible(out)

}
store(br.rmd2slides.pdf)

br.rmd2slides <- function(rmd.file,  theme = "AnnArbor", highlight = "tango", quiet = TRUE) {
### Make an R Markdown file into a Beamer PDF and and a Google Slides self-contained HTML file
### You need the proper BloomR edition

    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")

    html.file=paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.html')
    pdf.file=paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.pdf')

    br.rmd2slides.html(rmd.file, html.file, quiet = quiet)
    br.rmd2slides.pdf (rmd.file, pdf.file,  theme = theme, highlight = highlight, quiet = quiet)

}
store(br.rmd2slides)

highlight.styles <- function() {# List Pandoc syntax highlight styles 

    panexe <- .br.home("pandoc/bin/pandoc.exe")
    if(!file.exists(panexe)) stop("This function is not available in this BloomR edition")
    system2(panexe, "--list-highlight-styles")
}
store(highlight.styles)


```


br.rmd2pdf{#br.rmd2pdf} 
=======================

Description
-----------
Make an R markdown file into a PDF.

Usage
-----
    br.rmd2pdf(rmd.file, pdf.file, quiet=TRUE)

Arguments
---------
rmd.file
:   path to the R markdown file to be converted.  

pdf.file
:   path to the PDF file to be generated. If missing, change extension of rmd.file to pdf.  

quiet
:   FALSE to show the system PATH variable and intermediate files.  

Details
-------
You need BloomR LaTeX addons or the proper BloomR version. 

Value
-----
If there are no errors, it returns invisibly the absolute path of the output file.

```{r br.rmd2pdf, opts.label='purlme'}
br.rmd2pdf <- function(rmd.file, pdf.file, quiet=TRUE){
### Make an R Markdown file into a PDF
### You need BloomR LaTeX addons or the proper BloomR version
    
    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    if(missing(pdf.file)) pdf.file <- paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.pdf')
    if(pdf.file != basename(pdf.file)) stop("Sorry, currently you can only specify a name without directory as an output doc.")
    
    library(knitr)
    #library(rmarkdown)
    
    ## Render Rmd to PDF
    cmdexpr <- quote(
        rmarkdown::render(rmd.file,
               output_format=rmarkdown::pdf_document(highlight="tango",
                                          md_extensions="-tex_math_single_backslash"),
               output_file = pdf.file,
               clean = quiet, quiet = quiet)
    )
    out <- .br.pathexe(cmdexpr, quiet = quiet)
    invisible(out)
    
}
store(br.rmd2pdf)
```

br.rmd2both{#br.rmd2both} 
=========================

Description
-----------
Make an R Markdown file into a PDF and an HTML self-contained file

Usage
-----
    br.rmd2both(rmd.file, out.dir, quiet=TRUE)

Arguments
---------
rmd.file
:   path to the R markdown file to be converted.  

out.dir
:   directory of the output files. If missing, use `dirname(rmd.file)`.  

quiet
:   FALSE to show the system PATH variable and intermediate files.  

Details
-------
You need BloomR LaTeX addons or the proper BloomR version. 

Value
-----
If there are no errors, it returns invisibly the absolute path of the output file.


```{r br.rmd2both, opts.label='purlme'}
br.rmd2both <- function(rmd.file, quiet=TRUE){
### Make an R Markdown file into a PDF and an HTML self-contained file
### You need  BloomR LaTeX addons or the proper BloomR version

    
    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    
    pdf.file=paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.pdf')
    html.file=paste0(tools:::file_path_sans_ext(basename(rmd.file)), '.html')

    ## render does not support directory, perhaps we can move docs at the end. 
    ##if(!missing(out.dir)){
    ##    out.dir <- paste0(sub("/$", "", out.dir), "/") 
    ##    pdf.file <- paste0(out.dir, basename(pdf.file))
    ##    html.file <- paste0(out.dir, basename(html.file))
    ##}
    
    library(knitr)
    #library(rmarkdown)

    ## Render Rmd to HTML
    cmdexpr <- quote(
        rmarkdown::render(rmd.file,
               output_format=rmarkdown::html_document(theme="cerulean", highlight="tango",
                                           md_extensions="-tex_math_single_backslash"),
               output_file = html.file,
               clean = quiet, quiet = quiet)
    )
    out <- .br.pathexe(cmdexpr, pandonly = TRUE, quiet = quiet)

    if(is.na(out)) stop("HTML generation failed.")

    ## Render Rmd to PDF
    cmdexpr <- quote(
        rmarkdown::render(rmd.file,
               output_format = rmarkdown::pdf_document(highlight="tango",
                                            md_extensions = "-tex_math_single_backslash"),
               output_file = pdf.file,
               clean = quiet, quiet = quiet)
    )

    out <- .br.pathexe(cmdexpr, quiet = quiet)
    invisible(out)
}
store(br.rmd2both)

```


```{r to.be.finished, purl = FALSE, eval = FALSE, include = FALSE, echo = FALSE}

br.tex2pdf <- function(tex.file, pdf.file, auxname="latexaux", quiet=TRUE){

#### to be finished uncomment set path var lines, consider synctex !!!!!!!!! 

    
### Make an R Markdown file into a PDF
### You need the proper BloomR version

  
  ## Test arguments
  if(missing(tex.file)) stop("Argument 'tex.file' missing.")
  texsans <- tools:::file_path_sans_ext(tex.file)
  tex <- if(texsans == tex.file) paste0(texsans, '.tex') else tex.file
  pdf <- if(missing(pdf.file)) paste0(texsans, '.pdf') else pdf.file

  if(!file.exists(tex)) stop("There is no file\n", tex)
  if(!dir.exists(dirname(pdf))) stop("There is no directory\n", dirname(pdf))
  if(basename(auxname)!= auxname) stop("Please, use a name not a path as 'auxname' argument\n", auxname)
  
  ## Build Windows paths 
  pdfdir <- normalizePath(dirname(pdf))
  pdfbase <- basename(pdf)
  pdfbase.sq <- shQuote(tools:::file_path_sans_ext(pdfbase))
  texdir <- normalizePath(dirname(tex))
  texbase <- basename(tex)
  texbase.q <- shQuote(texbase)
  outdir <- auxname
  outdir.q <- shQuote(outdir)
  pdf.aux <- file.path(texdir, outdir, pdfbase)
    
  ## Set executable paths and build PDF
##############  old.path <- .br.addpaths(quiet = quiet)
  old.wd <- setwd(texdir)
  cmd <- c("pdflatex", "-interaction=batchmode",
           paste0("-output-directory=", outdir.q), paste0("-jobname=", pdfbase.sq), texbase.q)
  
  tryCatch(
    ret <- system2(cmd[1], cmd[-1]),
    finally = {
      setwd(old.wd)
      new.path <- Sys.getenv("Path")
##############      Sys.setenv(Path=old.path)
      if(ret) {
        cmd[2]  <- "" # remove batchmode
        stop("There was a non zero exit. To debug you can use:\n\n", 
             paste("path", new.path, "\n\n"),
             paste("cd", texdir, "\n\n"),
             paste(cmd, collapse = " "))
      } else   file.copy(pdf.aux, pdf)                         
        })
    invisible(ret)
}


```


Misc functions{#misc.functions}
==============================

Description
------------
`rm.all` deletes all objects (variables and functions) from memory, including invisible objects (those starting with a dot).
`rm.var` deletes non-function objects from memory.


Usage
-----
	rm.all()
	rm.var()
	

```{r miscfunc, opts.label='purlme'}

#Clean up
## Remove visible and invisible objects
rm.all <- function() 
    rm(list=ls(all=TRUE, envir=parent.frame()), envir=parent.frame())

## Remove visible non-function objects
rm.var <- function() 
    rm(list=setdiff(ls(envir=parent.frame()), lsf.str(envir=parent.frame())),  envir=parent.frame())

store(rm.all)
store(rm.var)
```

Beta functionalities{#beta.functions}
=====================================

Description
------------
Activate beta functionalities, if available for this release. 

Usage
-----
    br.beta()


```{r betafun, opts.label='purlme'}

br.beta <- function(){
    f=paste0(R.home("share"), "/bloomr/bloomr.beta.R")    
    if(file.exists(f)) source(f)  else message("No beta functionalities in this release")
}

store(br.beta)

```

```{r topics, purl=FALSE, eval=TRUE, include=FALSE}
read.head <- function(rmdfile) {x=readLines(rmdfile); x[grep("^=+", x)-1]}
topics <- function(rmdfile, mdfile){
    x <- read.head(rmdfile)
    x=sub("\\{", "]\\(", x )
    x=sub("\\}", "\\)", x )
    idx=paste0("[", x, "   ")
    idx=c("\nR topics documented:", "-----------", idx)
    txt <- readLines(mdfile)
    yaml.end=grep("^---", txt)[2] 
    txt=c(txt[1:yaml.end], idx, txt[(yaml.end+1):length(txt)])
    writeLines(txt, mdfile)
}
```

```{r parseargs, purl=FALSE, eval=TRUE, include=FALSE}


parseargs <- function(rmdfile, tmpfile) {
## Generate a new tmp Rmd with BloomR function arguments descriptions.
##    These are embedded in comments preceding (or in line with) the arguments and detected via a magic prefix tag
## The magic tag is hash (#) followed by a quote ('), henceforth denoted as HQ. In fact, to avoid side effects,
##    I can't explicitly write the magic tag. 
## Description comments have the format "HQ comment...". They can be in line with or precede the related arguments.
##    Standalone desc comments (i.e. non-inline) should immediately precede their argument, i.e. no-blank lines.
##    To break long lines, you can stack standalone comments above the argument, again without blank lines in the middle.
## Once extracted from comments, the actual markdown description is written when a special placeholder is met.
##    The placeholder is "HQ @args". You are expected to insert it in the markdown following the function and 
##    it will be replaced with the descriptions extracted from the comments in the function definition
##    HQ tag is always followed by a space (not a tab). Pay attention to casual use to avoid side effects. 
## Here is an example (replace HQ tag with hash and quote): 
##    
##    foo
##    ===
##    An important function.
##     
##    Usage
##    ------
##        foo(x,y)
##     
##      
##    Arguments
##    ----------
##     
##    HQ @args
##       
##    ```{r foo OPTIONS}      
##    foo <- function(
##                    HQ This is an important argument, and
##                    HQ it deserves special attention.
##                    x,
##                    y  HQ That's important too.
##                    ){
##     
##        BODY
##    }
##    ```

    
    txt=readLines(rmdfile)
    HQ=paste0("#", "'")

    ## Outer HQ comments
    ## -----------------
    ocrex.loose=sprintf("^ *%s +", HQ)   #-> " *HQ +"
    ocrex=sprintf("%s+[^@]", ocrex.loose) #-> " *HQ ++[^@]", ++ need perl

    ## Test standard outer comments in the middle of HQ comments
    x= grep("^ *#[^']", txt)
    if(length(x)) {
        if(x[1]==1) x=x[-1]
        if(x[length(x)]==length(txt)) x=x[-length(x)]
        xx=lapply(x, function(xx) {
            if(grepl(ocrex, txt[xx-1], , perl=TRUE) &  grepl(ocrex, txt[xx+1], perl=TRUE)) {        
                stop("Detected standard outer comments in the middle of HQ comments.\n",
                     paste(txt[xx+-1:1], collapse="\n" ))}    
        })
    }
    
    ## Get outer arg comments
    ocomp= grep(ocrex, txt, perl=TRUE)
    ocom=txt[ocomp]
    ## debug 
    ## ocom=letters[1:9]
    ## ocomp=c(1, 3:6, 10, 21:23)

    ## Identify contiguos outer comments
    ccom=c(TRUE, diff(ocomp)!=1)

    ## Stack all to first of contiguos comments 
    ocomp2=NULL; ocom2=NULL
    for(i in seq_along(ocomp)) {
        if(ccom[i]) {
            pos=ocomp[i]
            ocomp2=c(ocomp2, pos)
            val=sub(ocrex.loose, '', ocom[i])
            ocom2=c(ocom2, val)
        } else {
            val=paste(val, sub(ocrex.loose, '', ocom[i]))
            ocom2[length(ocom2)]=val        
        }    
    }
    ocomp=ocomp2; ocom=ocom2

    ##  Get args lines after comments
    findAgs=function(pos) { # find non comment line postion, given outer HQ comment position 
        while(startsWith(trimws(txt[pos]), "#")) pos=pos+1
        pos
    }      
    ocomp.ags=unlist(sapply(ocomp, findAgs))
    ocom.ags=txt[ocomp.ags]

    ## Test blanks after HQ comments
    ocom.ags=trimws(ocom.ags)
    if(any(sapply(ocom.ags, nchar) == 0))
        stop("Some HQ outer comments are followed by blank lines")
              
    ## Parse args line  
    end=regexpr("[=,)]|$", ocom.ags)
    ocom.ags=substr(ocom.ags, 1, end-1) 

    if(is.null(ocom)) ocom <- ocomp <- character(0)
    
    ## Inner HQ comments
    ## -----------------
    icrex.loose=sprintf("[^ ]+ *%s ", HQ)   #-> "[^ ]+ *HQ "

    ## Get inner arg comments
    icomp= grep(icrex.loose, txt)
    icom=txt[icomp]

    ## Extract args
    end=regexpr("[=,#]", icom)
    icom.ags=substr(icom, 1, end-1) 
    icom.ags=trimws(icom.ags)
    icom=sub(paste0(".+", HQ, " +"), '', icom)


    ## Match @args tag
    ## ---------------
    acrex=sprintf(" *%s +@args", HQ)   #-> " *HQ +@args"
    
    ## Get @args positions
    acomp= grep(acrex, txt)

    ## Merge argument slots and HQ comments by position
    df <- data.frame

    if(length(ocom)) ocom <- paste0(ocom.ags, "\n:   ", ocom)
    if(length(icom)) icom <- paste0(icom.ags, "\n:   ", icom)
    m <- merge(
        df(pos=ocomp, com=ocom),
        df(pos=icomp, com=icom),
        suffixes=c('.out','.in'),
        by="pos", all=TRUE)

    apos <- if(length(acomp)) TRUE else acomp
    m <- merge(
        df(pos=acomp, apos=apos),
        m,
        by="pos", all=TRUE)
    
    m[2]=!is.na(m[2])
    m[is.na(m)]=""
    m=df(m[1:2], com=apply(m[3:4], 1, paste, collapse=""))

    ## Split by comment by slots  and cat
    spl=rep(0, nrow(m))
    spl[m$apos] =  m$pos[m$apos]
    spl=cumsum(spl)
    spl=split(m$com, spl)
    spl=sapply(spl, function(x) paste(x[-1], collapse="\n\n"))


    ## Replace commment slots with comments 
    for(i in seq_along(acomp)){
        txt[acomp[i]] = spl[i]    
    }
    writeLines(txt,  tmpfile)
  
}
    
```

```{r attach, opts.label='purlme'}
### Make visible br.* in bloomr env and base ns
attach(bloomr.rmd)
rm(store)
rm(bloomr.rmd)

```

   
<!-- Local Variables: -->
<!-- mode: rmd -->
<!-- End: -->

<!--  LocalWords:  BloomR
 -->
