---
title: "BloomR Time Functions"
author: "Antonio Fasano"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
---

```{r build, purl=FALSE, eval=FALSE, include=FALSE}
THISLIB <- "bloomr.time"
## Polymode tends to start R synced with git root.
source(if(file.exists("br-common.R")) "br-common.R" else "src/br-libs/br-common.R")
knit.twice("bloomr-time.Rmd")

```

```{r CUSTOMIZE, purl=FALSE, include=FALSE}
## Resource directory relative path (not used here)
RESDIR <- "../../res"
```

```{r setup, purl=FALSE, include=FALSE}
eval(doc.setup()) # see br-common.R
```

```{r store, code = store.parse(THISLIB), opts.label='brfuncs'}
## Purl this first!
## Store br.* objects in dedicated namespace
## see br-common.R for store.parse()
```

# BloomR time functions


Beta time functions{#beta.misc.functions}
=========================================

Description
------------

Miscellaneous functions dealing with dates. 


Usage
-----
    br.try.date(d)
    br.is.same.class(...)


Arguments
---------
d
:   a POSIXlt, POSIXct, Date, "%Y/%m/%d", or "%Y-%m-%d" vector

Details
-------

`br.try.date` converts a vector to a date vector if possible or return `NULL`. Any vector element should be POSIXlt, POSIXct, Date, "%Y/%m/%d", or "%Y-%m-%d"

`br.is.same.class` check if all supplied argumets have the same class. It is mostly intended to check if dates are homogeneous. 


```{r MISCFUNC, opts.label='brfuncs'}

br.try.date=function(d){ # convert vector d to a date vector if possible or return null
### Any element should be POSIXlt, POSIXct, Date, "%Y/%m/%d", or "%Y-%m-%d"
                           
    if(identical(class(d), c("POSIXlt", "POSIXt"))) return(d) 
    if(identical(class(d), c("POSIXct", "POSIXt"))) return(d)
    if(identical(class(d), "Date")) return(d)

    if(is.character(d)) {
        
        try=as.Date(d, "%Y%m%d")
        if(!anyNA(try)) return(try)

        try=as.Date(d, "%Y/%m/%d")
        if(!anyNA(try)) return(try)

        try=as.Date(d, "%Y-%m-%d")
        if(!anyNA(try)) return(try)
    }
    
    return(NULL)
}

br.is.same.class=function(...){ # Check if all argumets have the same class
### Mostly intended to check if dates are homogeneous

    L=lapply(list(...), class)
    length(unique(L))==1
}


store(br.try.date)
store(br.is.same.class)


```


Time extension functions{#time.functions}
=========================================

Description
------------
Functions to get, set dates.

Usage
-----
    day(d)
    month(d)
    year(d)
    day(d, n)
    month(d, n)
    year(d, n)
    day(d)=x
    month(d)=x
    year(d)=x
    d %+% n
    d %-% n
    last.day(d)
    day.us(d1, d2)

Arguments
---------
d, d1, d2
:   objects of class date  

x
:   an integer representing the day/month/year  

n
:   an integer representing the months to add/subtract


Details
-------
If `component` is `day`, `month` or `year`: `component(d)` returns the *component* of the date `d` as an integer; `component(d, n)` returns the date `d` with the *component* set to the integer `n`; `component(d)= n` sets to the *component* of the date `d` to the integer `n`.  
`%+%` and `%-%` add and subtract months to a date.  
`last.day` returns last day of the month as an integer. `day.us` calculates date differences with the US convention.  


```{r time, opts.label='brfuncs'}
`%+%` <- function(x,y) UseMethod("%+%")
`%+%.Date` <- function(date,n) seq(date, by = paste (n, "months"), length = 2)[2]
`%-%` <- function(x,y) UseMethod("%-%")
`%-%.Date` <- function(date,n) seq(date, by = paste (-n, "months"), length = 2)[2]

year=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%Y")) else year(d)=n
    d    
}
`year<-`=function (d, value) d <-as.Date(paste0(value, format(d, "-%m-%d")))

month=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%m")) else month(d)=n
    d 
}
`month<-`=function (d, value) d <-as.Date(paste0(format(d, "%Y-"),  value, format(d, "-%d")))

day=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%d")) else day(d)=n
    d 
}
`day<-`=function (d, value) d <-as.Date(paste0(format(d, "%Y-%m-"),  value))

last.day=function(d){
    x=d %+% 1 #add a month
    day(x)=1  #set to 1st
    day(x-1)  #get day before
}

day.us=function(d1, d2){
    #set to first of month
    x1=day(d1,1);x2=day(d2,1);
    x=seq(x1, x2, by="1 month")
    #last day of each month in seq
    x=sapply(x, last.day)
    #count 31d-months
    x=length(which(x>30))
    #substract 1 for each 31d-month
    as.numeric(d2-d1-x)
}

store(day)
store(month)
store(year)
store(`day<-`)
store(`month<-`)
store(`year<-`)
store(`%+%`)
store(`%-%`)
store(`%+%.Date`, reg = TRUE)
store(`%-%.Date`, reg = TRUE)
store(last.day)
store(day.us)

```



```{r topics, purl=FALSE, eval=TRUE, include=FALSE}

# was for topics() and parseargs()

    
```

```{r attach, code = lib.attach(), opts.label='brfuncs'}
### Make visible br.* in bloomr env and base ns
## attach(get(THISLIB), name = THISLIB)
## rm(store)
## rm(list=c(THISLIB, "THISLIB"))

```
 

    
<!-- Local Variables: -->
<!-- mode: rmd -->
<!-- End: -->

<!--  LocalWords:  BloomR
 -->
