

# BloomR main functions


```{r setup, purl=FALSE, include=FALSE}
##rm(list=ls(all=TRUE))

### Shortcuts
opts_chunk$set(echo=TRUE)
OTS=opts_template$set
## Build function first
OTS(purlme=   list(purl=TRUE,  eval=!step2,   include=FALSE))
## Demos running after function building  
OTS(demofull= list(purl=FALSE, eval=step2,  include=TRUE))  # run w/ code + out   
OTS(demorun=  list(purl=FALSE, eval=step2,  include=TRUE, echo=FALSE)) # run w/ out, no code 
OTS(demodry=  list(purl=FALSE, eval=step2,  include=FALSE)) # run only (no code/out)
OTS(democode= list(purl=FALSE, eval=FALSE,  include=TRUE))  # show code only, no run  
OTS(buildenv= list(purl=FALSE, eval=step2,  include=FALSE))  # Called in step2 by the build chunk to knit this document 

## Purl hook
hook_purl=function (before, options, envir) 
{
    .knitEnv=knitr:::.knitEnv
    #isFALSE=knitr:::isFALSE
    isFALSE <- function (x) identical(x, FALSE)
    label_code=knitr:::label_code
    
    if (before || !options$purl || options$engine != "R") 
        return()
    output = .knitEnv$tangle.file
    if (isFALSE(.knitEnv$tangle.start)) {
        .knitEnv$tangle.start = TRUE
        unlink(output)
    }
    code = options$code
    #if (isFALSE(options$eval)) 
       ## code = comment_out(code, "# ", newline = FALSE)
    if (is.character(output)) {
        cat(label_code(code, options$params.src), file = output, 
            sep = "\n", append = TRUE)
    }
}
knit_hooks$set(purl = hook_purl)

```


```{r store, opts.label='purlme'}
## Purl this first
## Store br.* objects in bloomr env in base namespace
assign('bloomr',  new.env(parent=asNamespace("base")), envir=asNamespace("base"))

## func: store(func);  var: store("var")
store=function(sym){
    if(is.function(sym)){
        name=deparse(substitute(sym)) 
        val=sym
    } else {
        name=sym
        val=get(sym)
    }
    
    assign(name, val, envir=bloomr)
    rm( list=name, envir=parent.frame())
}

```



br.bdh{#br.bdh}
===============
*Historical data*

Description
-----------
Download historical Bloomberg data 

Usage
-----
    br.bdh(con, securities, fields="PX_LAST", start.date, end.date = NULL,
    option.names = NULL, option.values = NULL,
    always.display.tickers = FALSE, dates.as.row.names = (length(securities) == 1),
    include.non.trading.days = NULL
    ) 

Arguments
---------
con
:   the connection token returned from br.open()

securities
:   character vector of the tickers queried for data  

field
:   case insensitive character vector of the Bloomberg field queried. Defaults to "PX_LAST". 

start.date
:   Time series start date as a Date object or an ISO string without separators (YYYYMMDD). Time series will actually begin at `start.date` if there is data available; otherwise it will start at the first significant date. 

end.date
:   Time series end  date as a Date object or an ISO string without separators (YYYYMMDD). If NULL or missing, it defaults to the last available date. 

option.names, option.values
:   See details

always.display.tickers
:   Displays tickers in first column even if a single one is requested. Defaults to FALSE

dates.as.row.names  
:   Displays dates _also_ as row names. Defaults to TRUE for single ticker query, FALSE otherwise. 

include.non.trading.days
:   If TRUE, rows are returned for all requested dates, even when markets no data is available.
It defaults to FALSE for a single security or and TRUE for multiple securities. In the latter case, use `na.omit`  or `na.exclude` to remove these rows.


Details
-------
For multi-ticker queries you might consider to use `br.bulk.tiks` which features  also a simulated mode.

`option.names` and `option.values` are options vectors affecting the returned data. Options are set pairwise, for example to set `opt1` and `opt2`  respectively to `val1` and `val2`, you would pass the arguments:

    option.names=c("opt1", "opt2"), option.values=c("val1", "val2")

Here is a list of options:



periodicityAdjustment
:   Determine the frequency and calendar type of the output. To be used in conjunction with `periodicitySelection`. If `ACTUAL`, it  reverts to the actual date from today (if the end date is left blank) 
or from the End Date. If `CALENDAR`, (for pricing fields), it  reverts to the last business day of the specified calendar period. Calendar Quarterly (CQ), Calendar Semi-Annually (CS), or Calendar Yearly (CY). If `FISCAL`, it reverts to the fiscal period end for the company - Fiscal Quarterly (FQ), Fiscal Semi-Annually (FS) and Fiscal Yearly (FY) only.

periodicitySelection
:   Determine the frequency of the output. To be used in conjunction with periodicityAdjustment.
if `DAILY`, `WEEKLY`, `MONTHLY`, `QUARTERLY`, `SEMI_ANNUALLY`, `YEARLY`.

currency
:   Amends the value from local to desired currency. The value is a 3 letter ISO code string, e.g. USD, GBP. View `WCV<GO>` on the Bloomberg terminal for the full list.

overrideOption
:   Indicates whether to use the average or the closing price in quote calculation. Values can be 
`OVERRIDE_OPTION_CLOSE` for using the closing price or `OVERRIDE_OPTION_GPA`  for the average price.

pricingOption
:   Sets quote to Price or Yield for a debt instrument whose default value is quoted in yield (depending on pricing source).
`PRICING_OPTION_PRICE` sets quote to price; `PRICING_OPTION_YIELD` sets quote to yield.

nonTradingDayFillOption 
:   Sets to include/exclude non trading days where no data was generated.
`NON_TRADING_WEEKDAYS` includes all weekdays (Monday to Friday); `ALL_CALENDAR_DAYS`  includes all days of the calendar; `ACTIVE_DAYS_ONLY`  includes only the days where the instrument and field pair 
were updated.


nonTradingDayFillMethod
:   If data is to be displayed for non trading days what is the data to be returned.
 `PREVIOUS_VALUE` searches back and retrieve the previous value available for this security field pair. The search back period is up to one month. `NIL_VALUE` returns blank for the "value" value 
within the data element for this field.

maxDataPoints
:   the maximum number of data points to return. If the original data set is larger, the response will be a subset, containing only the last `maxDataPoints` data points.

returnEids
:   returns the entitlement identifiers associated with security. If TRUE, populates data with an extra element containing a name and value for the EID date.

returnRelativeDate
:   returns data with a relative date. If TRUE, populates data with an extra element containing a name and value for the relative date. For example RELATIVE_DATE = 2002 Q2

adjustmentNormal
:   Adjust for "change on day". If TRUE, adjusts historical pricing to reflect: Regular Cash, Interim, 1st Interim, 2nd Interim, 3rd Interim, 4th Interim, 5th Interim, Income, Estimated, Partnership Distribution, Final, Interest on Capital, Distribution, Prorated.

adjustmentAbnormal 
:   Adjusts for Anormal Cash Dividends. If TRUE, adjusts historical pricing to reflect: Special Cash, Liquidation, Capital Gains, Long-Term Capital Gains, Short-Term Capital Gains, Memorial, Return of Capital, Rights Redemption, Miscellaneous, Return Premium, Preferred Rights Redemption, Proceeds/Rights, Proceeds/Shares, Proceeds/Warrants.


adjustmentSplit
:   Capital Changes Defaults. If TRUE, adjusts historical pricing and/or volume to reflect: Spin-Offs, Stock Splits/Consolidations, Stock Dividend/Bonus, Rights Offerings/Entitlement.

adjustmentFollowDPDF
:   If TRUE (defaults) Follow the Bloomberg function as from `DPDF<GO>`.

CalendarCodeOverride
:   Returns the data based on the calendar of the specified country, exchange, or 
religion. Value is  a two character calendar code as from `CDR<GO>`. This will cause the data to be aligned according to the calendar  and including calendar holidays. Only applies only to DAILY requests.

calendarOverridesInfo
:   (Experimental, not tested) Returns data based on the calendar code of multiple countries, exchanges, or religious calendars as from `CDR<GO>`. This will cause the data to be aligned according to the set calendar(s) including their calendar holidays and only applies to DAILY requests.
Requires `calendarOverrides`, which is a character vector of  two-character calendar codes as from `CDR<GO>`; `calendareOverridesOperation`, which can be  `CDR_AND`  returning  the intersection of trading days among multiple calendars or  `CDR_OR` returning the union of trading days. That is, a data point is returned if a date is a valid trading day for any of the calendar codes specified in the request.

Overrides
:   (Experimental, not tested) Append overrides to modify the calculation. `fieldID` specifies  a field mnemonic or alpha-numeric, such as `PR092` or `PRICING_SOURCE`. Review FLDS<GO> for list of possible overrides. `value` sets the desired override value



Value
-----
A data frame with historical data. If tickers are displayed, the first column shows tickers, the second one the time series dates and the following ones the values of the queried fields; otherwise the columns start with dates. Dates will also be shown as rows if `dates.as.row.names=TRUE`. If multiple tickers are queried, they are vertically stacked respecting the order in `securities` vector.



```{r br.bdh, opts.label='purlme'}
br.bdh=function(
    con, securities, fields="PX_LAST", start.date, end.date = NULL,
    option.names = NULL, option.values = NULL,
    always.display.tickers = FALSE, dates.as.row.names = (length(securities) == 1),
    include.non.trading.days = NULL
    ) {
    if(is.null(con)) stop("'con' is NULL, but br.bdh() does not support simulated mode.")
    bdh(con, securities, fields, start.date, end.date, override_fields  = NULL, override_values = NULL,
        option.names, option.values, always.display.tickers, dates.as.row.names,
        include.non.trading.days)
}
store(br.bdh)
```


br.bulk.csv{#br.bulk.csv}
========================
*Historical data from grouped tickers in a CSV files*  
Reads a CSV file containing a group of tickers in each column and returns the historical data in xts or list format. The CSV file is assumed to have headers denoting group labels. 


Usage
------
    br.bulk.csv(con, file, start = Sys.Date() - 5, field = "PX_LAST", 
        cols = NULL, addtype = FALSE, showtype = FALSE, use.xts = TRUE, 
        comma = TRUE,
	    price=TRUE, nrow=5, same.dates=FALSE, no.na=FALSE, empty.sec=0
		) 
  
Arguments
----------

con
:   the connection token returned from br.open(). If `NULL` simulated values are generated.   

file
:   path to CSV file.  

start
:   start date. Can be a Date object or an ISO string without separators (YYYYMMDD). Defaults to 5 days before current date.  

field
:   case insensitive string denoting the Bloomberg field queried. Defaults to "PX_LAST". If the field is wrong or not accessible, data will be empty but no error will be raised.  

cols
:   Logical or integer vector for selecting CSV columns (ticker groups). Defaults to all columns.  

addtype
:   If a string denoting the security type, it will be added to all tickers; if TRUE "Equity", will be added; if FALSE (the default), nothing will be added.  

showtype
:   if TRUE, security types will be removed from names of list or xts output. It defaults to FALSE.  

use.xts
:   if TRUE (the default) each group will be formatted as an xts object else as a list.  

comma
:   to be set to FALSE for (non-English) CSV, using semicolon as separator.  

nrow
:   maximum number of simulated rows (actual is random). Ignored if `con!=NULL`, it defaults to 5.  

empty.sec
:   ratio of securities returning no data. Ignored if `con!=NULL`, it defaults to 0.

Details
-------
Empty CSV cells or cells interpreted as NAs will be ignored.  
If `con=NULL`, values are simulated by means of `br.sample()`. This function is used with default values, except for `nrow, start, same.dates, no.na, empty.sec`, which can be explicitly passed as arguments, and `sec.names` depending on tickers found in the CSV file. These arguments are ignored if `con!=NULL`. See `br.sample()` help for more.

Value
-----
a list where each element is the historical data of a CSV group.  
If `use.xts=TRUE`, elements are xts object, where each column is the historical data of a security.  
If `use.xts=FALSE`, elements are sub-list, where each element is the historical data of a security.  
If there is only one group, the first (and unique) element of the list will be returned.  


```{r br.bulk.csv, opts.label='purlme'}
br.bulk.csv=function(con, file, start=Sys.Date()-5, field="PX_LAST", cols=NULL,
    addtype=FALSE, showtype=FALSE, use.xts=TRUE, comma=TRUE,
    price=TRUE, nrow=5, same.dates=FALSE, no.na=FALSE, empty.sec=0
    )
{

    ## Check csv file
    if(!file.exists(file)) stop(paste('Unable to find the csv file', file))
    if(comma) csv=read.csv(file=file, as.is=TRUE) else{
        csv=read.csv2(file=file, as.is=TRUE)}

    ## Check cols arg
    if(!is.null(cols)){
        if(is.logical(cols) && !length(cols)==ncol(csv))
            stop(paste('Length of logical vector', paste(cols, collapse=' '),
                       'not equal to number of groups in', file))

        if(is.integer(cols) && max(cols)>ncol(csv))
            stop(paste('Unable to subset groups in', file, 'with columns', paste(cols, collapse=' ')))

        if(!is.logical(cols) && !all(cols%%1==0)) stop(paste(
         "'Col argument should be an integer or a logical vector of the same length of the groups in", file))
        csv=csv[cols]        
    }

    ## Get group names and count
    gnams=names(csv)
    gcnt=ncol(csv)

    ## Loop groups in csv
    grps=list()
    for(g in 1:gcnt){
        message('Processing ', gnams[g],  ' ...')
        x=list(br.bulk.tiks(con, csv[[g]],
            start, field, addtype, showtype, use.xts,
            price=price, nrow=nrow, same.dates=FALSE, no.na=FALSE, empty.sec=empty.sec))
        names(x)=gnams[g]
        grps=c(grps, x)        
    }
    if(length(grps)==1) grps=grps[[1]]
    grps
}
store(br.bulk.csv)
```

Demonstration
-------------


A sample CSV with Bloomberg tickers will look like follows:


```{r br.bulk.csv.demo1.show, opts.label='democode'}
read.csv("mybloomr/tickers.csv")
## This file is part of BloomR and anyway available here:
## https://github.com/AntonioFasano/BloomR/blob/master/res/tickers.csv
```

```{r br.bulk.csv.demo1, opts.label='demorun'}
read.csv("../res/tickers.csv")
```


Note: 

* CSV group headers are mandatory;
* Group headers need not to be the same length. 

We can now download data:

```{r br.bulk.csv.demo2, opts.label='demofull'}
con=NULL  #Simulated mode: replace with con=br.open() on terminal
```

```{r br.bulk.csv.demo3.show, opts.label='democode'}
data=br.bulk.csv(con, "mybloomr/tickers.csv") 
```

```{r br.bulk.csv.demo3, opts.label='demorun'}
data=br.bulk.csv(con, "../res/tickers.csv")
```


Above you see some info about data being processed that we will not show anymore in the following.

If you want to have detailed ticker descriptions,  see [br.bulk.desc Example](#br.bulk.desc.exem). Downloaded data look like follows:

```{r br.bulk.csv.demo4, opts.label='demofull'}
data
```

Note:

* The name of the securities tickers is stored without the security type: "Equity", "Index", etc.  
If this piece of info is significant for you, pass `showtype = TRUE`.   

* Time series start date defaults to 5 days before current date, unless you set `start` to: 
an R Date object (`start=as.Date("2014/9/30")`) or to a  more friendly ISO string (`start="20140930")`).     

Data are stored as a list of xts objects, each representing one group of tickers in the CSV file.


```{r br.bulk.csv.demo5, opts.label='demofull'}
length(data)
names(data)
class(data$Financial)
```

If you prefer you may get time series as data frames, and precisely as a list representing the ticker groups, where each group is in turn a list containing a data frame for each security:

```{r br.bulk.csv.demo6.q, opts.label='demodry'}
data=br.bulk.csv(con, "../res/tickers.csv", use.xts=FALSE) 
```
```{r br.bulk.csv.demo6.show, opts.label='democode'}
data=br.bulk.csv(con, "mybloomr/tickers.csv", use.xts=FALSE) 
```

```{r br.bulk.csv.demo7, opts.label='demofull'}
length(data)
names(data)
class(data$Financial)
length(data$Financial)
names(data$Financial)
class(data$Financial$`BAC US`)
```

By defaults time series list values from the Bloomberg "PX_LAST" field. To change the default field use:

```{r br.bulk.csv.demo8.show, opts.label='democode'}
data=br.bulk.csv(con, "mybloomr/tickers.csv", field = "PX_OPEN") 
```


You can choose to import only some of the CSV groups 

```{r br.bulk.csv.demo9, opts.label='demorun'}
data=br.bulk.csv(con, "../res/tickers.csv", cols=c(1,3))
```

```{r br.bulk.csv.demo9.show, opts.label='democode'}
data=br.bulk.csv(con, "mybloomr/tickers.csv", cols=c(1,3))
## or equivalently:
data=br.bulk.csv(con, "mybloomr/tickers.csv", cols=c(TRUE, FALSE, TRUE))
```

```{r br.bulk.csv.demo10, opts.label='demofull'}
names(data)
```
 
In the CSV file, if your tickers represent all equities, you can omit the type.   

Consider this CSV:

```{r br.bulk.csv.demo11.show, opts.label='democode'}
read.csv("mybloomr/tickers.eqt.csv")
## This file is part of BloomR and anyway available here:
## https://github.com/AntonioFasano/BloomR/blob/master/res/tickers.eqt.csv
```

```{r br.bulk.csv.demo11, opts.label='demorun'}
read.csv("../res/tickers.eqt.csv")
```

Note how the "Equity" type is missing! But you can use this CSV file with `addtype`:

```{r br.bulk.csv.demo12.q, opts.label='demodry'}
data=br.bulk.csv(con, "../res/tickers.eqt.csv", addtype=TRUE)
```

```{r br.bulk.csv.demo12.show, opts.label='democode'}
data=br.bulk.csv(con, "mybloomr/tickers.eqt.csv", addtype=TRUE)
```

Before going home, don't forget to:

```{r br.bulk.csv.demo13, opts.label='demofull'}
br.close(con)
```


br.bulk.desc{#br.bulk.desc}
===========================

Description
-----------
Get security descriptions for a vector of tickers.

Usage
-----
    br.bulk.desc(con, tiks) 

Arguments
---------
con
:    the connection token returned from br.open()  

tiks
:    character vector of the tickers queried for data  

Value
-----
A list of data frames, each representing the description of a security. For the format of data frames see the function `br.desc`.


```{r br.bulk.desc, opts.label='purlme'}
br.bulk.desc=function(con, tiks) {

    LL = lapply(tiks, function(tik){
        message('Reading ', tik)
        br.desc(con, tik)             
    })
    names(LL)=tiks
    LL
}
store(br.bulk.desc)
```

Example{#br.bulk.desc.exam}
----------------------------

```{r br.bulk.desc.exam.show,  opts.label='democode'}
con=br.open()
data=read.csv("mybloomr/tickers.csv", as.is=TRUE)
br.bulk.desc(con, as.vector(as.matrix(data[1:2,])))
br.close(con)
```


br.bulk.idx{#br.bulk.idx}
========================

Description
-----------
Returns the historical data for the constituents of an index in xts or list format.

Usage
-----

	br.bulk.idx(con, index, start = Sys.Date() - 5, field = "PX_LAST",  
	     showtype = FALSE, include.idx = TRUE, use.xts = TRUE, nsec = 50,  
         price = TRUE, nrow = 5, same.dates=FALSE, no.na=FALSE, empty.sec = 0,  
		 sec.names = NULL)  

Arguments
---------
con
:   the connection token returned from br.open(). If `NULL` simulated values are generated.   

index
:   string denoting the index ticker with or without the final security type label ('Index')  

include.idx
:   if TRUE (default) returns also historical data for the index.  

nsec
:   number of simulated index constituents. Ignored if `con!=NULL`, it defaults to 10.  

sec.names
:   character vector with names of sampled index constituents. Ignored if `con!=NULL`. By default security names are like 'memb1', 'memb2', etc.

For other arguments see the function `br.bulk.csv`

Details
-------
If `con=NULL`,  values are simulated by means of `br.sample()`. This function is used with default values, except for `nrow, nsec1, price, start, same.dates, no.na, empty.sec, sec.names`.

Value
-----
If `use.xts=TRUE`, an xts object, where each column is the historical data of a constituent.  
If `use.xts=FALSE`, a list, where each element is the historical data of a constituent.  
If `include.idx=TRUE`, the last column or element will be the historical data of the index.  


```{r br.bulk.idx, opts.label='purlme'}
br.bulk.idx=function(con, index, start=Sys.Date()-5, field="PX_LAST", showtype=FALSE,
    include.idx=TRUE, use.xts=TRUE,
    nsec=10, price=TRUE, nrow=5,
    same.dates=FALSE, no.na=FALSE, empty.sec=0, sec.names = NULL
    )
{
   
    ## Check connection
    if(!is.null(con) && !.br.is.con(con)) stop('Invalid connection parameter')

    ## Check index format. Add 'INDEX' if missing
    if(!is.character(index)) stop('Index should be a string')
    if(length(index)>1) stop('Only one index')
    if(!grepl("INDEX$", toupper(index))) index=paste(index, 'INDEX') 

    ## Get index members
    if(is.null(con)) tiks=paste0('memb', 1:nsec) else{
        tiks=bds(con, index, 'INDX_MEMBERS')
        tiks=paste(tiks[[1]], 'Equity')
    }

    ## Check sec.names
    if(is.null(con) && !is.null(sec.names)) {
        if(!is.character(sec.names)) stop("'sec.names' should be a character vector")
        if(length(sec.names)!=nsec) stop("'sec.names' length should be equal to the number of index constituents")
        tiks=sec.names
    }
    
    ## Include index?
    if(include.idx) tiks=c(tiks, index)

    ## Get data
    br.bulk.tiks(
        con=con, 
        tiks=tiks, 
        start=start,
        field=field,
        addtype=FALSE,
        showtype=showtype,
        use.xts=use.xts,
        price=price,
        nrow=nrow,
        same.dates=same.dates,
        no.na=no.na,
        empty.sec=empty.sec
        )
}
store(br.bulk.idx)
```


br.bulk.tiks{#br.bulk.tiks}
==========================
*Bulk historical data*  
Returns the historical data for a vector of tickers in xts or list format

Usage
------
    br.bulk.tiks(con, tiks, start=Sys.Date()-5, field="PX_LAST",
		addtype=FALSE, showtype=FALSE, use.xts=TRUE,
		price=TRUE, nrow=5, same.dates=FALSE, no.na=FALSE, empty.sec=0) 
  
Arguments
---------
tiks
:   character vector of the tickers queried for data  

For other arguments see the function `br.bulk.csv` 

Details
--------
If an element of `tiks` is `NA` or empty (`""`) it is ignored. This is intended to avoid errors when the character vector are read from a CSV file with empty cells.  
If `con=NULL`, values are simulated by means of `br.sample()`. Sampled values are based on default values of `br.sample()`, but it is possible to set explicitly  `start, same.dates, no.na, empty.sec`; `sec.names` depends on `tiks` argument. These arguments are ignored if `con!=NULL`. See `br.sample()` help for more.

Value
------
If `use.xts=TRUE`, an xts object, where each column is the historical data of a security.  
If `use.xts=FALSE`, a list, where each element is the historical data of a security.



Example
-------

```{r br.bulk.tiks.exam, opts.label='demofull'}
con=NULL # Open simulated connection and  load some data
br.bulk.tiks(con, c("MSFT US", "AMZN US"), addtype=TRUE)
br.close(con) # Use the token to release the connection
```

See Also
--------

[br.bulk.csv](#br.bulk.csv)

```{r br.bulk.tiks, opts.label='purlme'}
br.bulk.tiks=function(
    con,  
    tiks, 
    start=Sys.Date()-5, field="PX_LAST",
    addtype=FALSE, showtype=FALSE, use.xts=TRUE,
    price=TRUE, nrow=5, same.dates=FALSE, no.na=FALSE, empty.sec=0
    )
{ 

    ## Check connection
    if(!is.null(con) && !.br.is.con(con)) stop('Invalid connection parameter') 
    
    ## Check tickers (skip possible empty CSV cells)
    if(!is.character(tiks)) stop('Tickers should be in the form of a character vector')
    tiks=tiks[!is.na(tiks)] 
    tiks=tiks[tiks!='']
    
    ## Check start date
    if(is.na(as.Date(start, format='%Y%m%d'))) stop(paste('Invalida date', start))

    ## Check security type to add/show
    .br.check.type(addtype)
    .br.check.type(showtype)
    if(addtype==TRUE)  addtype="Equity"
    if(addtype!=FALSE) tiks=paste(tiks, addtype)
    if(!showtype) tiks.show=.br.cuttype(tiks) else tiks.show=tiks
    
    ## Check xts library availability
    if(use.xts && !require("xts", quietly=TRUE, character.only=TRUE))  stop("Can't find library xts")

    ## Get data as an xts class
    if(use.xts){ 
        LL = lapply(tiks, function(tik){
            message('Loading ', tik)
            if(!is.null(con)) x=br.bdh(con, tik, field, start) else {
                x=br.sample(nrow, 1, price=price, start=start,
                    df=TRUE, same.dates=same.dates, no.na=no.na,
                    sec.names=c('date', field))
            }
            x=xts(x[-1], as.Date (x[[1]]))
            if(nrow(x)==0) x=NA else x
        })

        ## Randomly identify empty.sec
        if(is.null(con)){
            x=round(length(tiks.show) * empty.sec) 
            empty=sample(length(tiks.show), x)
            LL[empty] = NA 
        }

        ##print(LL ) ###################
        ## If there are only NAs, output an empty xts else merge ts and in simul. mode use no.na val  
        if(all(is.na(LL))) LL=na.omit(xts(t(unlist(LL)), Sys.Date())) else {
          #  if(!is.null(con)) no.na=FALSE # in merge c(LL, all=!no.na)
            if(length(LL)>1) LL=do.call("merge.xts", LL) else LL=LL[[1]]
        }
    
        ## Set labels
        if(nrow(LL)>0) names(LL) = tiks.show  else  dimnames(LL)= list(NULL,tiks.show)
        LL
        
    } else {
        ## Get data in list format
        LL=lapply(tiks, function(tik){
            message('Loading ', tik)
            if(!is.null(con)) br.bdh(con, tik, field, start) else {
                br.sample(nrow, 1, price=price, empty.sec=empty.sec, start=start,
                           df=TRUE, sec.names=c('date', field))
            }            
        })
        
        ## Randomly identify empty.sec
        if(is.null(con)){
            x=round(length(tiks.show) * empty.sec) 
            empty=sample(length(tiks.show), x)
            LL[empty] = NA 
        }
        
        setNames(LL, tiks.show)
    }
}
store(br.bulk.tiks)
```

br.desc{#br.desc}
================

Description
-----------
Get security descriptions.

Usage
-----
    br.desc(con, tik)

Arguments
---------
con
:   the connection token returned from br.open()  

tik
:   string denoting the ticker queried for data  

Value
-----
A data frame containing the value of the Bloomberg fields form `ds001` to `ds009` and the long field `CIE_DES_BULK`.

```{r br.desc, opts.label='purlme'}
br.desc=function(con, tik)
{

    ## Check connection
    if(!.br.is.con(con)) stop('Invalid connection parameter') 

    ## Check ticker format
    if(!is.character(tik)) stop('The ticker should be a string')
    if(length(tik)>1) stop('Only one  ticker')
       
    ## Short description fields as data frame
    des=paste0('ds00', 1:9)
    des=des[-7] # not usually working 
    x=bdp(con, tik, des)
    x=data.frame(t(x), stringsAsFactors=FALSE)

    ## Long description field
    xx=bds(con, tik, 'CIE_DES_BULK')

    ## Merge fields add long desc to DF
    if(!is.null(xx)) colnames(xx) = colnames(x)
    rnams=c(rownames(x), rownames(xx))
    x=rbind(x,xx)
    rownames(x)=rnams
    x
}
store(br.desc)
```


```{r rmd-internal, opts.label='purlme'}

.br.addpaths <- function(pandonly = FALSE, quiet=TRUE){
### Add to Windows System Path the executable directories of LaTeX, Pandoc, and Perl with this search priority, and
### return invisibly the original path. If "pandonly" is true, add only Pandoc. If "quiet" is false, print the new path.

    ## Exit in case this snippet is not exectuted in Windows
    if(.Platform$OS.type != "windows")
        stop("Sorry, Bloomberg Terminal only exists for Windows and so BloomR functions.")
        
  ## Find executables
  panexe=dbr.brmain("pandoc/bin/pandoc.exe")
  if(!file.exists(panexe))
    stop(paste("Unable to find:", panexe, '\nYour BloomR edition might not support it.'))
  pandir=normalizePath(dirname(panexe))
    
  latbin=dbr.brmain("latex/texmfs/install/miktex/bin/x64/latex.exe")
  if(!file.exists(latbin) && !pandonly){
    stop(paste("Unable to find:", latbin, '\nYour BloomR edition might not support it.'))
    latdir=normalizePath(dirname(latbin))
  }

  perlexe <- dbr.brmain("perl/bin/perl.exe")
  if(!file.exists(perlexe) && !pandonly){
    stop(paste("Unable to find:", perlexe, '\nYour BloomR edition might not support it.'))
    perldir <- normalizePath(dirname(perlexe))
  }
    
  ## Add executable dirs to system path
  old.path=Sys.getenv('Path')
  f <- function(dir) Sys.setenv(PATH=paste0(dir, ";", Sys.getenv("PATH")))

  ## Wanted final order is: LaTeX;Pandoc;Perl 
  if(!pandonly) f(perldir)
  f(pandir)
  if(!pandonly) f(latdir)
  
  ## Add Ghostscript environment variable   
  if(!pandonly)  Sys.setenv(GSC="mgs.exe")   

  if(!quiet) message(Sys.getenv("PATH"))

  invisible(old.path)
  
}


```

br.md2pdf{#br.md2pdf} 
=====================

Description
-----------
Make a markdown file into a PDF
It assumes that you have installed the proper BloomR version.

Usage
-----
    br.md2pdf(md.file, pdf.file)

Arguments
---------
md.file
:   path to the markdown file to be converted.  

pdf.file
:   path to the PDF file to be generated.  

quiet
:   FALSE to show the system PATH variable.  

Details
-------
The function will stop with an error if you have not installed proper BloomR version.

Value
-----
If there are no errors, it returns zero invisibly, otherwise it prints an error message and returns the related error code.


```{r br.md2pdf, opts.label='purlme'}
br.md2pdf=function(md.file, pdf.file, quiet=FALSE){
### Make a markdown file into a PDF
### You need the proper BloomR version

    ## Test arguments
    if(missing(md.file)) stop("Argument 'md.file' missing.")
    if(missing(pdf.file)) stop("Argument 'pdf.file' missing.")
    
    ## Set executable paths and render
    old.path <- .br.addpaths(quiet=quiet)
    cmd <- paste("pandoc", .br.wpath(md.file), '-o', .br.wpath(pdf.file))
    tryCatch(
        out  <-  system(cmd, intern = TRUE, invisible = FALSE),
        finally = {
            ## Restore original system Path
            Sys.setenv(Path=old.path)
            ## Return errors if any
            if(!is.null(attr(out, 'status')))  message(paste(out, collapse="\n"))
        })  

    invisible(out)
}
store(br.md2pdf)
```

br.rmd2html{#br.rmd2html} 
========================

Description
-----------
Make an R markdown file into a HTML.

Usage
-----
    br.rmd2html(rmd.file, html.file, quiet=TRUE)

Arguments
---------
rmd.file
:   path to the R markdown file to be converted.  

html.file
:   path to the HTML file to be generated. If missing, change extension of rmd.file to html.  

quiet
:   FALSE to show the system PATH variable.  

Details
-------
You need BloomR LaTeX addons or the proper BloomR version. 

Value
-----
If there are no errors, it returns invisibly the absolute path of the output file.


```{r br.rmd2html, opts.label='purlme'}
br.rmd2html=function(rmd.file, html.file, quiet=TRUE){
### Make an R Markdown file into an HTML self-contained file
### You need  BloomR LaTeX addons or the proper BloomR version

    
    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    if(missing(html.file)) html.file=paste0(tools:::file_path_sans_ext(rmd.file), '.html')

    library(knitr)
    library(rmarkdown)

    ## Set executable paths and render
    old.path <- .br.addpaths(pandonly = TRUE, quiet = TRUE)  
    tryCatch(
        ## Render Rmd to HTML
        out <- render(rmd.file,
                      output_format=html_document(theme="cerulean", highlight="tango",
                                                  md_extensions="-tex_math_single_backslash"),
                      output_file=html.file, quiet=quiet),
        finally = {
            ## Restore original system Path
            Sys.setenv(Path=old.path)
        })

    invisible(out)
}
store(br.rmd2html)
```


br.rmd2pdf{#br.rmd2pdf} 
=======================

Description
-----------
Make an R markdown file into a PDF.

Usage
-----
    br.rmd2pdf(rmd.file, pdf.file, quiet=TRUE)

Arguments
---------
rmd.file
:   path to the R markdown file to be converted.  

pdf.file
:   path to the PDF file to be generated. If missing, change extension of rmd.file to pdf.  

quiet
:   FALSE to show the system PATH variable and intermediate files.  

Details
-------
You need BloomR LaTeX addons or the proper BloomR version. 

Value
-----
If there are no errors, it returns invisibly the absolute path of the output file.

```{r br.rmd2pdf, opts.label='purlme'}
br.rmd2pdf=function(rmd.file, pdf.file, quiet=TRUE){
### Make an R Markdown file into a PDF
### You need BloomR LaTeX addons or the proper BloomR version
    
    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    if(missing(pdf.file)) pdf.file=paste0(tools:::file_path_sans_ext(rmd.file), '.pdf')

    library(knitr)
    library(rmarkdown)
    
    ## Set executable paths and render
    old.path <- .br.addpaths(quiet = TRUE)    
    tryCatch(
        ## Render Rmd to PDF
        out <- render(rmd.file,
                      output_format=pdf_document(highlight="tango",
                                                  md_extensions="-tex_math_single_backslash"),
                      output_file=pdf.file,
                      clean=quiet, quiet=quiet),
        finally = {
            ## Restore original system Path
            Sys.setenv(Path=old.path)
        })
    invisible(out)            
}
store(br.rmd2pdf)
```

br.rmd2both{#br.rmd2both} 
=========================

Description
-----------
Make an R Markdown file into a PDF and an HTML self-contained file

Usage
-----
    br.rmd2both(rmd.file, out.dir, quiet=TRUE)

Arguments
---------
rmd.file
:   path to the R markdown file to be converted.  

out.dir
:   directory of the output files. If missing, use `dirname(rmd.file)`.  

quiet
:   FALSE to show the system PATH variable, and intermediate files.  

Details
-------
You need BloomR LaTeX addons or the proper BloomR version. 

Value
-----
If there are no errors, it returns invisibly the absolute path of the output file.


```{r br.rmd2both, opts.label='purlme'}
br.rmd2both=function(rmd.file, out.dir, quiet=TRUE){
### Make an R Markdown file into a PDF and an HTML self-contained file
### You need  BloomR LaTeX addons or the proper BloomR version

    
    ## Test arguments
    if(missing(rmd.file)) stop("Argument 'rmd.file' missing.")
    
    pdf.file=paste0(tools:::file_path_sans_ext(rmd.file), '.pdf')
    html.file=paste0(tools:::file_path_sans_ext(rmd.file), '.html')
    if(!missing(out.dir)){
        out.dir <- paste0(sub("/$", "", out.dir), "/") 
        pdf.file <- paste0(out.dir, basename(pdf.file))
        html.file <- paste0(out.dir, basename(html.file))
    }
    
    library(knitr)
    library(rmarkdown)

    ## Set executable paths
    old.path <- .br.addpaths(quiet = TRUE)
    
    tryCatch(
        {
            ## Render Rmd to HTML
            out <- render(rmd.file,
                          output_format=html_document(theme="cerulean", highlight="tango",
                                                     md_extensions="-tex_math_single_backslash"),
                          output_file=html.file, quiet=quiet)

            ## Render Rmd to PDF
            out <- render(rmd.file,
                          output_format=pdf_document(highlight="tango",
                                                     md_extensions="-tex_math_single_backslash"),
                          output_file=pdf.file,
                          clean=quiet, quiet=quiet)

        },
        finally = {
            ## Restore original system Path
            Sys.setenv(Path=old.path)
        })
    
    invisible(dirname(out))    
}
store(br.rmd2both)

```

```{to.be.finished}

br.tex2pdf=function(tex.file, pdf.file, auxname="latexaux", quiet=TRUE){
### Make an R Markdown file into a PDF
### You need the proper BloomR version

  ## to be finished uncomment set path var lines, consider synctex !!!!!!!!! 
  
  ## Test arguments
  if(missing(tex.file)) stop("Argument 'tex.file' missing.")
  texsans <- tools:::file_path_sans_ext(tex.file)
  tex <- if(texsans == tex.file) paste0(texsans, '.tex') else tex.file
  pdf <- if(missing(pdf.file)) paste0(texsans, '.pdf') else pdf.file

  if(!file.exists(tex)) stop("There is no file\n", tex)
  if(!dir.exists(dirname(pdf))) stop("There is no directory\n", dirname(pdf))
  if(basename(auxname)!= auxname) stop("Please, use a name not a path as 'auxname' argument\n", auxname)
  
  ## Build Windows paths 
  pdfdir <- normalizePath(dirname(pdf))
  pdfbase <- basename(pdf)
  pdfbase.sq <- shQuote(tools:::file_path_sans_ext(pdfbase))
  texdir <- normalizePath(dirname(tex))
  texbase <- basename(tex)
  texbase.q <- shQuote(texbase)
  outdir <- auxname
  outdir.q <- shQuote(outdir)
  pdf.aux <- file.path(texdir, outdir, pdfbase)

  library(knitr)
  library(rmarkdown)
    
  ## Set executable paths and build PDF
##############  old.path <- .br.addpaths(quiet = quiet)
  old.wd <- setwd(texdir)
  cmd <- c("pdflatex", "-interaction=batchmode",
           paste0("-output-directory=", outdir.q), paste0("-jobname=", pdfbase.sq), texbase.q)
  
  tryCatch(
    ret <- system2(cmd[1], cmd[-1]),
    finally = {
      setwd(old.wd)
      new.path <- Sys.getenv("Path")
##############      Sys.setenv(Path=old.path)
      if(ret) {
        cmd[2]  <- "" # remove batchmode
        stop("There was a non zero exit. To debug you can use:\n\n", 
             paste("path", new.path, "\n\n"),
             paste("cd", texdir, "\n\n"),
             paste(cmd, collapse = " "))
      } else   file.copy(pdf.aux, pdf)                         
        })
    invisible(ret)
}



```

br.sample{#br.sample}
====================

Description
------------
Return simulated historical data for n securities in xts or df format.

Usage
-----
    br.sample(nrow, nsec=1, price=TRUE, start=Sys.Date(),
	mean=ifelse(price, 10, 0.1), sd=1, jitter=0, same.dates=FALSE, no.na=FALSE,  
	empty.sec=0, df=FALSE, sec.names=NULL)

Arguments
---------
nrow
:   number of simulated data points for each security; if `same.dates=FALSE`, the number of rows for each sampled security will be a random number not exceeding nrow, else it will be nrow for all securities.  

nsec
:   number of simulated securities (defaults to 1).  

price
:   if TRUE (default), simulated values are non-negative.  

start
:   start date. Can be a Date object or an ISO string without separators (YYYYMMDD). Defaults to current date.  

mean
:   mean of security generated values. If `price=TRUE`, default to 10 else defaults to 0.1.  

sd
:   sd of security generated values. It defaults to 1.  

jitter
:   modifies each security mean by adding adding a random value in [-jitter, jitter]. Defaults to 0.  

same.dates
:   if TRUE, all sampled securities will refer to the same dates and for each security the number will equal nrow. If FALSE (default), date values and number will randomly differ. For each security the random number will not exceed `nrow`.  

no.na
:   if `same.dates=FALSE`, when merging sampled security data NAs are likely to be produced. If `no.na=FALSE` (default) they will be left, otherwise they will be removed using R `na.omit`  

df
:   if FALSE (default), the output will be an xts object, else the output will be a data frame with the first column containing the dates of the sampled data.  

sec.names
:   character vector for column names. If `df=FALSE` the length of the vector should be equal to `nsec`, else to `nsec + 1` (because of the first column containing dates). By default security names are like 'sample1', 'sample2', etc. and the date column is named 'date'.  

empty.sec
:   ratio of securities returning no data (defaults to 0). The result is rounded without decimal places.  

Value
-----
If `df=TRUE`, a data frame object, where the first column is the vector with all generated dates merged and each  subsequent column contains the sampled data of a security. If `df=FALSE`, an xts object, where each element is the sampled data of a security, while the dates will be part of the xts time object. In both cases if `same.dates=FALSE` and/or `empty.sec!=0` generated data points will have different length and the the date gaps will be filled with NAs, except if `no.na=TRUE`. If the generated values are only NAs the output will be converted to a 0-rows xts or data frame, containing only security labels accessible with `dimnames(*)[[2]]`. 

```{r br.sample, opts.label='purlme'}
br.sample=function(nrow, nsec=1, price=TRUE, start=Sys.Date(), mean=ifelse(price, 10, 0.1), sd=1,
    jitter=0, same.dates=FALSE, no.na=FALSE, df=FALSE, empty.sec=0, sec.names=NULL)
{
 
    if(!require("xts", quietly=TRUE, character.only=TRUE))  stop("Can't find library xts")

    ## Start can be Date class or ISO string without sep
    if(is.na(as.Date(start, format='%Y%m%d'))) stop(paste('Invalida date', start)) else 
        start=as.Date(start, '%Y%m%d')

    ## Check sec.names
    if(!is.null(sec.names)) {
        if(!is.character(sec.names)) stop("'sec.names' should be a character vector")
        if(!df && length(sec.names)!=nsec) stop("'sec.names' length should be equal to 'nsec'")
        if(df && length(sec.names)!=nsec+1) stop("'sec.names' length should be equal to 'nsec'+1")
    }

    ## Check empty.sec is a ratio
    if(empty.sec<0 || empty.sec>1) stop("'empty.sec' must be between 0 and 1")

    ## Randomly identify empty.sec
    x=round(nsec * empty.sec) 
    empty=sample(nsec, x)
    
    ## Make xts matrix    
    tss=lapply(1:nsec, function(col){

        ## Jitter 
        mean.jit= mean + runif(1, -jitter, jitter)
    
        ## Generate TS with fixed or random dates
        if(same.dates) r=nrow else r=sample(1:nrow,1)

        ## Generate column if not among the empty ones
        if(col %in% empty){
            x=NA
        } else {            
            x=xts(round(rnorm(r,mean.jit, sd),3), sort(sample(start+1:nrow-1,r)))            
            if(price) coredata(x)=abs(coredata(x))   # Price always non-negative
        }
                
        x
    })

    ## If there are only NAs cells output an empty date/val xts else merge and fill empty cells with NAs 
    if(all(is.na(tss))) tss=na.omit(xts(t(unlist(tss)), Sys.Date())) else {
        if(length(tss)>1) tss=do.call("merge.xts", tss) else
        tss=tss[[1]]
    }

    ## Remove NAs
    if(no.na) tss=na.omit(tss)
    
    ## Set labels
    x=paste0('sample', 1:ncol(tss))
    if(nrow(tss)>0)  names(tss) = x  else  dimnames(tss)= list(NULL,x)
        
    ## Convert to data frame
    if(df) tss=data.frame(date=time(tss), tss)
    
    if(is.null(sec.names)) tss else
       setNames(tss, sec.names)
   
}
store(br.sample)
```


Deprecated functions{#deprecated.functions}
===========================================

Description
------------
Functions not used anymore generating an informative error

Usage
-----
    bbg.open()
    bbg.close(con)
 
Arguments
---------
con
:   the connection token returned from br.open()

Example
-------

    con=bbg.open()
	##  Sorry 'bbg.open' is now deprecated. Please use br.open().

```{r deprecated, opts.label='purlme'}
bbg.open=function() stop("Sorry 'bbg.open' is now deprecated. Please use br.open().")
bbg.close=function(con) stop("Sorry 'bbg.close' is now deprecated. Please use br.close().")
store(bbg.open)
store(bbg.close)
```


Internal BloomR functions{#Internal}
=====================================

Description:
------------
Internal functions not to be used by the end user

Usage:
------

    .br.is.con(con)
    .br.types
    .br.check.type(type) 
    .br.cuttype(type)
    .br.jar()

Arguments:
----------
con
:   the connection token returned from br.open()  

type
:   a string representing the security type  

Details
-------
`.br.is.con` checks for the validity of a connection token.
`.br.types` is a character vector with security types suitable as an argument for `br.bulk*` functions.
`.br.check.type` checks if a type matches `.br.types`.
`.br.cuttype` cuts trailing security type from character vector.
`.br.jar()` returns the path to the blpapi*.jar


```{r bbg-internal, opts.label='purlme'}

## Check connection token
.br.is.con=function(con) identical(attr(con, 'jclass'), "org/findata/blpwrapper/Connection")

## Legal security types
.br.types=c('Govt', 'Corp', 'Mtge', 'M-Mkt', 'Muni', 'Pfd', 'Equity', 'Comdty', 'Index', 'Curncy')

## Check security type
.br.check.type=function(type) {
    if(is.character(type)){
	x=toupper(type)
	xx=toupper(.br.types)
	if(!any(xx %in% x)) stop(paste(x, 'not in', paste(xx, collapse=' ')))
    }
}

## Cut trailing security type from character vector 
.br.cuttype=function(type){
    p=paste0(' +', .br.types, '$|', collapse='')
    p=sub('\\|$', '', p)
    sub(p, '', type, ignore.case=TRUE)
}


.br.jar=function(){
    jarpath=dbr.brmain("/blpapi/bin")
    Sys.glob(file.path(jarpath,  "blpapi-[0-9]*.jar"))
    }


store(.br.is.con)
store(".br.types")
store(.br.check.type) 
store(.br.cuttype)
store(.br.jar)
```

Manage connections{#connections}
===============================

Description
------------
Open and close the connection to the Bloomberg service.   


Usage
-----
    br.open()
    br.close(con)
	
Arguments
---------
con
:   the connection token returned from br.open()

Details
-------

`br.open` returns the connection token needed by the BloomR function downloading data. When you finish you session, you pass it to `br.close`. If you are using simulated data and so your connection token is NULL, closing the connection is optional. Anyway running `br.close(con)`, even if `con==NULL` avoids adding this line when you switch to a actual data download.


Example
-------

```{r connections.exem, opts.label='democode'}
con=br.open() # Open the connection and get the token and load some data
br.bulk.tiks(con, c("MSFT US", "AMZN US"), addtype=TRUE)
br.close(con) # Use the token to release the connection

```


```{r connections, opts.label='purlme'}
br.open=function() blpConnect(blpapi.jar.file=.br.jar())
br.close=function(conn) if(!is.null(conn)) blpDisconnect(conn)

store(br.open)
store(br.close)
```


Misc functions{#misc.functions}
==============================

Description
------------
`rm.all` deletes all objects (variables and functions) from memory, including invisible objects (those starting with a dot).
`rm.var` deletes non-function objects from memory.


Usage
-----
	rm.all()
	rm.var()
	

```{r miscfunc, opts.label='purlme'}

#Clean up
## Remove visible and invisible objects
rm.all=function() 
    rm(list=ls(all=TRUE, envir=parent.frame()), envir=parent.frame())

## Remove visible non-function objects
rm.var=function() 
    rm(list=setdiff(ls(envir=parent.frame()), lsf.str(envir=parent.frame())),  envir=parent.frame())

store(rm.all)
store(rm.var)
```

Beta functionalities{#beta.functions}
=====================================

Description
------------
Activate beta functionalities, if available for this release. 

Usage
-----
    br.beta()


```{r betafun, opts.label='purlme'}

br.beta=function(){
    f=paste0(R.home("share"), "/bloomr/bloomr.beta.R")    
    if(file.exists(f)) source(f)  else message("No beta functionalities in this release")
}

store(br.beta)

```

Time extension functions{#time.functions}
=========================================

Description
------------
Functions to get, set dates.

Usage
-----
    day(d)
    month(d)
    year(d)
    day(d, n)
    month(d, n)
    year(d, n)
    day(d)=x
    month(d)=x
    year(d)=x
    d %+% n
    d %-% n
    last.day(d)
    day.us(d1, d2)

Arguments
---------
d, d1, d2
:   objects of class date  

x
:   an integer representing the day/month/year  

n
:   an integer representing the months to add/subtract


Details
-------
If `component` is `day`, `month` or `year`: `component(d)` returns the *component* of the date `d` as an integer; `component(d, n)` returns the date `d` with the *component* set to the integer `n`; `component(d)= n` sets to the *component* of the date `d` to the integer `n`.  
`%+%` and `%-%` add and subtract months to a date.  
`last.day` returns last day of the month as an integer. `day.us` calculates date differences with the US convention.  


```{r time, opts.label='purlme'}
`%+%` <- function(x,y) UseMethod("%+%")
`%+%.Date` <- function(date,n) seq(date, by = paste (n, "months"), length = 2)[2]
`%-%` <- function(x,y) UseMethod("%-%")
`%-%.Date` <- function(date,n) seq(date, by = paste (-n, "months"), length = 2)[2]

year=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%Y")) else year(d)=n
    d    
}
`year<-`=function (d, value) d <-as.Date(paste0(value, format(d, "-%m-%d")))

month=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%m")) else month(d)=n
    d 
}
`month<-`=function (d, value) d <-as.Date(paste0(format(d, "%Y-"),  value, format(d, "-%d")))

day=function(d, n=NULL){
    if(is.null(n)) d=as.numeric(format(d, "%d")) else day(d)=n
    d 
}
`day<-`=function (d, value) d <-as.Date(paste0(format(d, "%Y-%m-"),  value))

last.day=function(d){
    x=d %+% 1 #add a month
    day(x)=1  #set to 1st
    day(x-1)  #get day before
}

day.us=function(d1, d2){
    #set to first of month
    x1=day(d1,1);x2=day(d2,1);
    x=seq(x1, x2, by="1 month")
    #last day of each month in seq
    x=sapply(x, last.day)
    #count 31d-months
    x=length(which(x>30))
    #substract 1 for each 31d-month
    as.numeric(d2-d1-x)
}

store(day)
store(month)
store(year)
store(`day<-`)
store(`month<-`)
store(`year<-`)
store(`%+%`)
store(`%-%`)
store(`%+%.Date`)
store(`%-%.Date`)
store(last.day)
store(day.us)

```

```{r topics, purl=FALSE, eval=TRUE, include=FALSE}
read.head=function() {x=readLines('bloomr.Rmd'); x[grep("^=+", x)-1]}
topics=function(){
    x=read.head()
    x=sub("\\{", "]\\(", x )
    x=sub("\\}", "\\)", x )
    idx=paste0("[", x, "   ")
    idx=c("R topics documented:", "-----------", idx)
    txt=readLines('bloomr.md')
    first=grep(".", txt)[1]
    txt=c(txt[first], idx, txt[(first+1):length(txt)])
    writeLines(txt, 'bloomr.md')
#    c(txt[first], idx,  txt[(first+1):(first+4)])
}
```


```{r attach, opts.label='purlme'}
### Make visible br.* in bloomr env and base ns
attach(bloomr)
rm(store)
```

```{r setpaths, opts.label='buildenv'}

if(build.env$use.bloomr)
    original.system.path <- .br.addpaths()

```

```{r build, purl=FALSE, eval=FALSE, include=FALSE}
knit2=function(file, use.bloomr=TRUE){
### Knit this Rmd with BloomR shipped Pandoc & LaTeX, if use.bloomr=T and we're on Windows. 
    
    if(use.bloomr && .Platform$OS.type != "windows")
        stop("If you want to build this Rmd with BloomR shipped Pandoc & LaTeX, you need to use Windows.")
        
    require(knitr); require(rmarkdown)
    knitr:::opts_chunk$set(tidy.opts=list(width.cutoff=60))
    fse=tools:::file_path_sans_ext(file)
    step2=FALSE
    knit(file)
    step2=TRUE
    build.env <- new.env()
    build.env$use.bloomr <- use.bloomr
    knit(file, envir = build.env)
    if(use.bloomr && .Platform$OS.type != "windows")   
        message("Old Path: ", build.env$original.system.path, 
                "\nNew: ", Sys.getenv("PATH"))
    topics()
    render(paste0(fse, '.md'), "html_document")
    render(paste0(fse, '.md'), pdf_document())
    render("../README.md", html_document())

    if(use.bloomr && .Platform$OS.type != "windows"){
        Sys.setenv(Path=build.env$original.system.path)
        message("Old system path restored")
    }
}
## knit2("bloomr.Rmd", use.bloomr=FALSE) 
knit2("bloomr.Rmd") 

## if with markdown lib, but topics needs rewrite
## shell("pandoc bloomr.md -o bloomr.pdf", shell=Sys.getenv("COMSPEC"))
## markdownToHTML("bloomr.md", "bloomr.html") 
## markdownToHTML("README.md", "README.html")

```

```{r unitests, purl=FALSE, eval=FALSE, include=FALSE}
con=br.open() 
br.bulk.tiks(con, c("MSFT US", "AMZN US"), addtype=TRUE)
br.bulk.idx(con, "SX5E Index", field="PX_LAST")
br.close(con) 
```


    
<!-- Local Variables: -->
<!-- mode: rmd -->
<!-- End: -->

<!--  LocalWords:  BloomR
 -->
