
br.bdh{#br.bdh}
===============
*Historical data*

XXXXRemove always.display.ticker, dates.as.row.names?
XXXXCheck only.trading.days description
XXXXMove staff in auto args

Description
-----------
Download historical Bloomberg data

Usage
-----
    br.bdh(con, security, fields="PX_LAST", start.date=Sys.Date() - 7, end.date=Sys.Date(),
    override_fields= NULL, override_values=NULL,
    option.names = NULL, option.values = NULL,
    only.trading.days = TRUE
    )

Arguments
---------
con
:   the connection token returned from br.open()

securities
:   character vector of the tickers queried for data

field
:   case insensitive character vector of the Bloomberg field queried. Defaults to "PX_LAST".

start.date
:   Time series start date as a Date object or an ISO string without separators (YYYYMMDD). Time series will actually begin at `start.date` if there is data available; otherwise it will start at the first significant date.

end.date
:   Time series end  date as a Date object or an ISO string without separators (YYYYMMDD). If NULL or missing, it defaults to the last available date.

option.names, option.values
:   See details

always.display.tickers
:   (removed) Displays tickers in first column even if a single one is requested. Defaults to FALSE

dates.as.row.names
:   (removed) Displays dates _also_ as row names. Defaults to TRUE for single ticker query, FALSE otherwise.

only.trading.days
:   If FALSE, rows are returned for all requested dates, even when no markets data is available.
XXXXCheck! It defaults to FALSE for a single security or and TRUE for multiple securities. In the latter case, use `na.omit`  or `na.exclude` to remove these rows.


Details
-------
For multi-ticker queries you might consider to use `br.hist()` which features also a simulated mode.

`option.names` and `option.values` are options vectors affecting the returned data. Options are set pairwise, for example to set `opt1` and `opt2`  respectively to `val1` and `val2`, you would pass the arguments:

    option.names=c("opt1", "opt2"), option.values=c("val1", "val2")

Here is a list of options:



periodicityAdjustment
:   Determine the frequency and calendar type of the output. To be used in conjunction with `periodicitySelection`. If `ACTUAL`, it  reverts to the actual date from today (if the end date is left blank)
or from the End Date. If `CALENDAR`, (for pricing fields), it  reverts to the last business day of the specified calendar period. Calendar Quarterly (CQ), Calendar Semi-Annually (CS), or Calendar Yearly (CY). If `FISCAL`, it reverts to the fiscal period end for the company - Fiscal Quarterly (FQ), Fiscal Semi-Annually (FS) and Fiscal Yearly (FY) only.

periodicitySelection
:   Determine the frequency of the output. To be used in conjunction with periodicityAdjustment.
if `DAILY`, `WEEKLY`, `MONTHLY`, `QUARTERLY`, `SEMI_ANNUALLY`, `YEARLY`.

currency
:   Amends the value from local to desired currency. The value is a 3 letter ISO code string, e.g. USD, GBP. View `WCV<GO>` on the Bloomberg terminal for the full list.

overrideOption
:   Indicates whether to use the average or the closing price in quote calculation. Values can be
`OVERRIDE_OPTION_CLOSE` for using the closing price or `OVERRIDE_OPTION_GPA`  for the average price.

pricingOption
:   Sets quote to Price or Yield for a debt instrument whose default value is quoted in yield (depending on pricing source).
`PRICING_OPTION_PRICE` sets quote to price; `PRICING_OPTION_YIELD` sets quote to yield.

nonTradingDayFillOption
:   Sets to include/exclude non trading days where no data was generated.
`NON_TRADING_WEEKDAYS` includes all weekdays (Monday to Friday); `ALL_CALENDAR_DAYS`  includes all days of the calendar; `ACTIVE_DAYS_ONLY`  includes only the days where the instrument and field pair
were updated.


nonTradingDayFillMethod
:   If data is to be displayed for non trading days what is the data to be returned.
 `PREVIOUS_VALUE` searches back and retrieve the previous value available for this security field pair. The search back period is up to one month. `NIL_VALUE` returns blank for the "value" value
within the data element for this field.

maxDataPoints
:   the maximum number of data points to return. If the original data set is larger, the response will be a subset, containing only the last `maxDataPoints` data points.

returnEids
:   returns the entitlement identifiers associated with security. If TRUE, populates data with an extra element containing a name and value for the EID date.

returnRelativeDate
:   returns data with a relative date. If TRUE, populates data with an extra element containing a name and value for the relative date. For example RELATIVE_DATE = 2002 Q2

adjustmentNormal
:   Adjust for "change on day". If TRUE, adjusts historical pricing to reflect: Regular Cash, Interim, 1st Interim, 2nd Interim, 3rd Interim, 4th Interim, 5th Interim, Income, Estimated, Partnership Distribution, Final, Interest on Capital, Distribution, Prorated.

adjustmentAbnormal
:   Adjusts for Anormal Cash Dividends. If TRUE, adjusts historical pricing to reflect: Special Cash, Liquidation, Capital Gains, Long-Term Capital Gains, Short-Term Capital Gains, Memorial, Return of Capital, Rights Redemption, Miscellaneous, Return Premium, Preferred Rights Redemption, Proceeds/Rights, Proceeds/Shares, Proceeds/Warrants.


adjustmentSplit
:   Capital Changes Defaults. If TRUE, adjusts historical pricing and/or volume to reflect: Spin-Offs, Stock Splits/Consolidations, Stock Dividend/Bonus, Rights Offerings/Entitlement.

adjustmentFollowDPDF
:   If TRUE (defaults) Follow the Bloomberg function as from `DPDF<GO>`.

CalendarCodeOverride
:   Returns the data based on the calendar of the specified country, exchange, or
religion. Value is  a two character calendar code as from `CDR<GO>`. This will cause the data to be aligned according to the calendar  and including calendar holidays. Only applies only to DAILY requests.

calendarOverridesInfo
:   (Experimental, not tested) Returns data based on the calendar code of multiple countries, exchanges, or religious calendars as from `CDR<GO>`. This will cause the data to be aligned according to the set calendar(s) including their calendar holidays and only applies to DAILY requests.
Requires `calendarOverrides`, which is a character vector of  two-character calendar codes as from `CDR<GO>`; `calendareOverridesOperation`, which can be  `CDR_AND`  returning  the intersection of trading days among multiple calendars or  `CDR_OR` returning the union of trading days. That is, a data point is returned if a date is a valid trading day for any of the calendar codes specified in the request.

Overrides
:   (Experimental, not tested) Append overrides to modify the calculation. `fieldID` specifies  a field mnemonic or alpha-numeric, such as `PR092` or `PRICING_SOURCE`. Review FLDS<GO> for list of possible overrides. `value` sets the desired override value



Value
-----
A data frame with historical data. If tickers are displayed, the first column shows tickers, the second one the time series dates and the following ones the values of the queried fields; otherwise the columns start with dates. Dates will also be shown as rows if `dates.as.row.names=TRUE`. If multiple tickers are queried, they are vertically stacked respecting the order in `securities` vector.


```{r br.bdh, opts.label='brfuncs'}
br.bdh=function(
                con, security, fields="PX_LAST",
                start.date=Sys.Date() - 7, end.date=Sys.Date(),
                override_fields  = NULL, override_values = NULL,
                option.names = NULL, option.values = NULL,
                only.trading.days = TRUE
                ) {

    if(is.null(con)) stop("'con' is NULL, but br.bdh() does not support simulated mode.")


    ## Set overrides
    if(is.null(override_fields)) {
        override_fields="IGNORE"
        override_values = ""
    }

    ## Test/set dates
    dates=.br.test.dates(start.date, end.date, asChar=TRUE)
    option.names=c(option.names, "startDate")
    option.values=c(option.values, dates$start)
    if(!is.null(end.date)) {
        option.names = c(option.names, "endDate")
        option.values = c(option.values, dates$end)
    }


    ## Non trading days
    if(!only.trading.days) {
        option.names=c("nonTradingDayFillOption", "nonTradingDayFillMethod", option.names)
        option.values=c("ALL_CALENDAR_DAYS", "NIL_VALUE", option.values)
    }

    ## Get reference data
    ref=con$blh(security, .jarray(fields),
            .jarray(override_fields), .jarray(override_values),
            .jarray(option.names), .jarray(option.values))
    ret=ref$getData()
    ret=if(!is.null(ret)) .jevalArray(ret, simplify=TRUE) else ret

    ## Set data attributes
    if(!is.null(ret)) attr(ret, "types")=ref$getDataTypes()
    ## For info on data types Rbbg:::convert.to.type
    ## Consider also ref$getColumnNames()

    ret
}
store(br.bdh)
```


```{r moretests, purl=FALSE, eval=FALSE, include=FALSE}
# br.bdh(con, "MSFT US EQUITY", "PX_LAST", "20170202", "20170210", override_fields="PRICING_SOURCE", override_values="CG")
# br.bdh(con, "MSFT US EQUITY", "PX_LAST", "20170202", "20170210")

####
# debug(br.hist) ; br.hist(con, c("MSFT US Equity", "AMZN US Equity"), start="20170225")
#debug(br.hist.csv)
#debug(br.hist)
# br.hist.csv(NULL, "wei.csv", same.dates=TRUE, empty.sec=.2)

# br.bdh(con, "MSFT US Equity", start="20170225")

```

